---
title: "Statistical analysis using DESeq2"
author: "Hildah Njoroge"
date: "2023-07-25"
output: html_document
---

## Load the data from the HTG pipeline

```{r}
# load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/htg_pipeline/htg_dds_filtering.Rdata")
```

## Load the data from the HTSeq pipeline

```{r}
# load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/hisat2_htseq_pipeline/htseq_dds_filtering.Rdata")

load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/filtering_criteria.Rdata")
```

## Assumptions of a linear model (and why we have considered a generalised linear model)

1.  linearity of predictor and response variables , i.e., changes in these two variables are proportional and consistent
2.  Homoskedasticity, i.e., reponse variable variance is the same for all the mean values of the predoctor variables
3.  Normally distributed (reponse?)
4.  independent observations

```{r}
hist(counts(dds_rowSums_greaterthan_10_HTSeq), ylim = c(0,5000))
hist(counts(dds_cpm_filtered_HTSeq1), ylim = c(0,5000))
# hist(rowMeans(log2(cpm(counts(dds)) + 1)))
```
#### QQ plots of the data
```{r}
# hist(counts(dds_rowSums_greaterthan_10_HTSeq), ylim = c(0,5000))
qqnorm(counts(dds_cpm_filtered_HTSeq1), ylim = c(0,5000))
qqline(counts(dds_cpm_filtered_HTSeq1), ylim = c(0,5000))
# hist(rowMeans(log2(cpm(counts(dds)) + 1)))
```

## Attempting to draw QQ plot of the count data --\> FAILED

```{r}
# # Example data (replace this with your raw gene count data)
# # (counts_data <- matrix(rnbinom(1000, mu = 20, size = 2), ncol = 1))
# head(counts_data <- countTable)
# 
# # Calculate quantiles of the theoretical normal distribution
# theoretical_quantiles <- qnorm(ppoints(nrow(counts_data)))
# 
# # Calculate quantiles of the observed data
# observed_quantiles <- quantile(counts_data, ppoints(nrow(counts_data)))
# 
# # Create a data frame for the Q-Q plot
# qq_plot_data <- data.frame(Theoretical = theoretical_quantiles, Observed = observed_quantiles)
# 
# plot(qq_plot_data)
# 
# # Create the Q-Q plot using ggplot2
# # ggplot(qq_plot_data, aes(x = Theoretical, y = Observed)) +
# #   geom_point() +
# #   geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
# #   xlab("Theoretical Quantiles") +
# #   ylab("Observed Quantiles") +
# #   ggtitle("Q-Q Plot of Raw Gene Count Data")

```

## Attempt to draw mean vs standard deviation scatter plot

# Steps to statistical analysis usng DESeq2

## Step 1: Define design matrix

```{r}

# designMatrix <- model.matrix(~ 0 + disease + individual, data = sampleTable)
designMatrix <- model.matrix(~ 0 + condition, data = sampleTable)
head(designMatrix)
```

## Step 2: Define contrast matrix

### noColitis vs colitisOnly

```{r}
contrastMatrixCol <- makeContrasts(conditioncolitisOnly - conditionnoColitis, levels = designMatrix)

head(contrastMatrixCol)

```

### noColitis vs colitis_nodysplasia

```{r}
contrastMatrixColnodysp <- makeContrasts(conditioncolitis_nodysplasia - conditionnoColitis, levels = designMatrix)

head(contrastMatrixColnodysp)
```

### noColitis vs colitis_dysplasia

```{r}
contrastMatrixColdysp <- makeContrasts(conditioncolitis_dysplasia - conditionnoColitis, levels = designMatrix)

head(contrastMatrixColdysp)
```

### colitisOnly vs colitis_nodysplasia
```{r}
contrastMatrix2Colnodysp <- makeContrasts(conditioncolitis_nodysplasia - conditioncolitisOnly, levels = designMatrix)

head(contrastMatrix2Colnodysp)
```

### colitisOnly vs colitis_dysplasia
```{r}
contrastMatrix2Coldysp <- makeContrasts(conditioncolitis_dysplasia - conditioncolitisOnly, levels = designMatrix)

head(contrastMatrix2Coldysp)
```

### colitis_nodysplasia vs colitis_dysplasia
```{r}
contrastMatrix3Colnodysp <- makeContrasts(conditioncolitis_dysplasia - conditioncolitis_nodysplasia, levels = designMatrix)

head(contrastMatrix3Colnodysp)
```
```{r}
contrastMatrixAll <- makeContrasts(conditioncolitisOnly - conditionnoColitis, conditioncolitis_nodysplasia - conditionnoColitis, conditioncolitis_dysplasia - conditionnoColitis, levels = designMatrix)

head(contrastMatrixAll)
```

#Step 3: Fit model 
1. calculate the normalisation factors 
2. estimate the dispersion parameters 
3. fit the model to the data 

### HTSeq

```{r}

countTable <- as.data.frame(counts(dds_cpm_filtered_HTSeq1))
(dge <- DGEList(countTable))
(dge <- calcNormFactors(dge))
(dge <- estimateDisp(dge, designMatrix, robust = TRUE))
fit <- glmQLFit(dge, designMatrix, robust = TRUE)
```

## Step 4: Perform hypothesis testing

### 1. Perform hypothesis testing

```{r}
(res <- glmQLFTest(fit, contrast = contrastMatrixCol))
```

### 2. filter genes, i.e., (i) create a table of top DEG (Tags) sorted by Pvalue

```{r}
(res <- topTags(res, n = nrow(countTable)))
nrow(res)
```

### ... (ii) based on the level of significance (adjusted Pvalue, FDR adjusted Pvalue of 0.05) and the level of relevance (log2FC \> 1, i.e., FC of at least 2)

```{r}
sigRes <- subset(res$table, FDR < 0.05 & abs(logFC) > 1)
sigRes_HTSeq <- sigRes
knitr::kable(head(sigRes))
# htseq_tags <- rownames(res$table)[1:20]; htseq_tags
nrow(sigRes)
# rownames(sigRes)[1:20]
```

## To add a column of up- and down-regulated genes

```{r}
deg_reg <- data.frame("regulation" = NULL)

#to confirm the number of DEG is correct
nrow(sigRes_HTSeq); sum(nrow(subset(sigRes_HTSeq,sigRes_HTSeq$logFC > 1)),
nrow(subset(sigRes_HTSeq,sigRes_HTSeq$logFC < 1)))


for (i in sigRes_HTSeq$logFC) {
  # print(i)
  if (i > 1) {
    new_row <- data.frame(regulation = "up")
    deg_reg <- rbind(deg_reg,new_row)
  } else {new_row <- data.frame(regulation = "down")
    deg_reg <- rbind(deg_reg,new_row)}
  }

sigRes_HTSeq_reg <- cbind(sigRes_HTSeq, deg_reg)
nrow(sigRes_HTSeq_reg)
```


## Visualize results in a Volcano plot

```{r}
library(ggplot2)
library(ggrepel)

volcanoPlot <- ggplot(res$table,
                      aes(x = logFC, y = -log10(FDR),
                          color = ifelse(FDR < 0.05 & abs(logFC) > 1, "darkred", "grey"))) +
  geom_point() +
  xlab(expression("Fold Change, Log"[2]*"")) +
  ylab(expression("Adjusted P value, Log"[10]*"")) +
  geom_vline(xintercept = c(-1, 1), linetype = "dotted", linewidth = 1) +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", linewidth = 1) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_colour_manual(values = c("darkred", "grey", "steelblue")) +
  geom_text_repel(aes(x = logFC, y = -log10(FDR), label = rownames(res$table)[1:10],
                      size = 2, color = "steelblue"),
                  data = res$table[1:10, ])
print(volcanoPlot)
```


# B. noColitis vs colitis_nodysplasia
## Step 1: Define design matrix
## Step 2: Define contrast matrix
### noColitis vs colitis_nodysplasia

```{r}
contrastMatrixColnodysp <- makeContrasts(conditioncolitis_nodysplasia - conditionnoColitis, levels = designMatrix)

head(contrastMatrixColnodysp)
```

#Step 3: Fit model 1. calculate the normalisation factors 2. estimate the dispersion parameters 3. fit the model to the data 
#Step 3: Fit model
1. calculate the normalisation factors
2. estimate the dispersion parameters
3. fit the model to the data

### HTSeq
```{r}

countTable <- as.data.frame(counts(dds_cpm_filtered_HTSeq1))
(dge <- DGEList(countTable))
(dge <- calcNormFactors(dge))
(dge <- estimateDisp(dge, designMatrix, robust = TRUE))
fit <- glmQLFit(dge, designMatrix, robust = TRUE)
```

## Step 4: Perform hypothesis testing

### 1. Perform hypothesis testing

```{r}
(res <- glmQLFTest(fit, contrast = contrastMatrixColnodysp))
nrow(res)
```

### 2. filter genes, i.e., (i) create a table of top DEG (Tags) sorted by Pvalue

```{r}
(res <- topTags(res, n = nrow(countTable)))
nrow(res)
```

### ... (ii) based on the level of significance (adjusted Pvalue, FDR adjusted Pvalue of 0.05) and the level of relevance (log2FC \> 1, i.e., FC of at least 2)

```{r}
sigRes <- subset(res$table, FDR < 0.05 & abs(logFC) > 1)
sigRes_HTSeq_Colnodysp <- sigRes
knitr::kable(head(sigRes))
# htseq_tags <- rownames(res$table)[1:20]; htseq_tags
nrow(sigRes)
# rownames(sigRes)[1:20]
```

## To add a column of up- and down-regulated genes

```{r}
deg_reg <- data.frame("regulation" = NULL)

#to confirm the number of DEG is correct
nrow(sigRes_HTSeq_Colnodysp) ; sum(nrow(subset(sigRes_HTSeq_Colnodysp,sigRes_HTSeq_Colnodysp$logFC > 1)),nrow(subset(sigRes_HTSeq_Colnodysp,sigRes_HTSeq_Colnodysp$logFC < 1)))


for (i in sigRes_HTSeq_Colnodysp$logFC) {
  # print(i)
  if (i > 1) {
    new_row <- data.frame(regulation = "up")
    deg_reg <- rbind(deg_reg,new_row)
  } else {new_row <- data.frame(regulation = "down")
    deg_reg <- rbind(deg_reg,new_row)}
  }

sigRes_HTSeq_Colnodysp_reg <- cbind(sigRes_HTSeq_Colnodysp, deg_reg)
```


## Visualize results in a Volcano plot

```{r}
volcanoPlot <- ggplot(res$table,
                      aes(x = logFC, y = -log10(FDR),
                          color = ifelse(FDR < 0.05 & abs(logFC) > 1, "darkred", "grey"))) +
  geom_point() +
  xlab(expression("Fold Change, Log"[2]*"")) +
  ylab(expression("Adjusted P value, Log"[10]*"")) +
  geom_vline(xintercept = c(-1, 1), linetype = "dotted", linewidth = 1) +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", linewidth = 1) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_colour_manual(values = c("darkred", "grey", "steelblue")) +
  geom_text_repel(aes(x = logFC, y = -log10(FDR), label = rownames(res$table)[1:10],
                      size = 2, color = "steelblue"),
                  data = res$table[1:10, ])
print(volcanoPlot)
```

# C. noColitis vs colitis_dysplasia
## Step 1: Define design matrix
## Step 2: Define contrast matrix


### noColitis vs colitis_dysplasia

```{r}
contrastMatrixColdysp <- makeContrasts(conditioncolitis_dysplasia - conditionnoColitis, levels = designMatrix)

head(contrastMatrixColdysp)
```


#Step 3: Fit model 1. calculate the normalisation factors 2. estimate the dispersion parameters 3. fit the model to the data 
#Step 3: Fit model
1. calculate the normalisation factors
2. estimate the dispersion parameters
3. fit the model to the data

### HTSeq
```{r}

countTable <- as.data.frame(counts(dds_cpm_filtered_HTSeq1))
(dge <- DGEList(countTable))
(dge <- calcNormFactors(dge))
(dge <- estimateDisp(dge, designMatrix, robust = TRUE))
fit <- glmQLFit(dge, designMatrix, robust = TRUE)
```

## Step 4: Perform hypothesis testing

### 1. Perform hypothesis testing

```{r}
(res <- glmQLFTest(fit, contrast = contrastMatrixColdysp))
```

### 2. filter genes, i.e., (i) create a table of top DEG (Tags) sorted by Pvalue

```{r}
(res <- topTags(res, n = nrow(countTable)))
nrow(res)
```

### ... (ii) based on the level of significance (adjusted Pvalue, FDR adjusted Pvalue of 0.05) and the level of relevance (log2FC \> 1, i.e., FC of at least 2)

```{r}
sigRes <- subset(res$table, FDR < 0.05 & abs(logFC) > 1)
sigRes_HTSeq_Coldysp <- sigRes
knitr::kable(head(sigRes))
# htseq_tags <- rownames(res$table)[1:20]; htseq_tags
nrow(sigRes)
# rownames(sigRes)[1:20]
```


## To add a column of up- and down-regulated genes

```{r}
deg_reg <- data.frame("regulation" = NULL)

#to confirm the number of DEG is correct
nrow(sigRes_HTSeq_Coldysp) ; sum(nrow(subset(sigRes_HTSeq_Coldysp,sigRes_HTSeq_Coldysp$logFC > 1)),nrow(subset(sigRes_HTSeq_Coldysp,sigRes_HTSeq_Coldysp$logFC < 1)))


for (i in sigRes_HTSeq_Coldysp$logFC) {
  # print(i)
  if (i > 1) {
    new_row <- data.frame(regulation = "up")
    deg_reg <- rbind(deg_reg,new_row)
  } else {new_row <- data.frame(regulation = "down")
    deg_reg <- rbind(deg_reg,new_row)}
  }

sigRes_HTSeq_Coldysp_reg <- cbind(sigRes_HTSeq_Coldysp, deg_reg)
```

## Visualize results in a Volcano plot

```{r}
volcanoPlot <- ggplot(res$table,
                      aes(x = logFC, y = -log10(FDR),
                          color = ifelse(FDR < 0.05 & abs(logFC) > 1, "darkred", "grey"))) +
  geom_point() +
  xlab(expression("Fold Change, Log"[2]*"")) +
  ylab(expression("Adjusted P value, Log"[10]*"")) +
  geom_vline(xintercept = c(-1, 1), linetype = "dotted", linewidth = 1) +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", linewidth = 1) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_colour_manual(values = c("darkred", "grey", "steelblue")) +
  geom_text_repel(aes(x = logFC, y = -log10(FDR), label = rownames(res$table)[1:10],
                      size = 2, color = "steelblue"),
                  data = res$table[1:10, ])
print(volcanoPlot)
```

====== 
# Identfy DEG among the diseased groups
#### contrastMatrix2Colnodysp
#### contrastMatrix2Coldysp
#### contrastMatrix3Colnodysp

#Step 3: Fit model 
1. calculate the normalisation factors 
2. estimate the dispersion parameters 
3. fit the model to the data 

### HTSeq

```{r}

countTable <- as.data.frame(counts(dds_cpm_filtered_HTSeq1))
(dge <- DGEList(countTable))
(dge <- calcNormFactors(dge))
(dge <- estimateDisp(dge, designMatrix, robust = TRUE))
fit <- glmQLFit(dge, designMatrix, robust = TRUE)
```

## Step 4: Perform hypothesis testing

### 1. Perform hypothesis testing

### Colitisonly vs colnodysp

```{r}
(res <- glmQLFTest(fit, contrast = contrastMatrix2Colnodysp))
nrow(res)
```

### 2. filter genes, i.e., (i) create a table of top DEG (Tags) sorted by Pvalue

```{r}
(res <- topTags(res, n = nrow(countTable)))
nrow(res)
```

### ... (ii) based on the level of significance (adjusted Pvalue, FDR adjusted Pvalue of 0.05) and the level of relevance (log2FC \> 1, i.e., FC of at least 2)

```{r}
sigRes <- subset(res$table, FDR < 0.05 & abs(logFC) > 1)
sigRes_HTSeq_2_Colnodysp <- sigRes
knitr::kable(head(sigRes))
# htseq_tags <- rownames(res$table)[1:20]; htseq_tags
nrow(sigRes)
# rownames(sigRes)[1:20]
```



## To add a column of up- and down-regulated genes

```{r}
deg_reg <- data.frame("regulation" = NULL)

#to confirm the number of DEG is correct
nrow(sigRes_HTSeq_2_Colnodysp) ; sum(nrow(subset(sigRes_HTSeq_2_Colnodysp,sigRes_HTSeq_2_Colnodysp$logFC > 1)),nrow(subset(sigRes_HTSeq_2_Colnodysp,sigRes_HTSeq_2_Colnodysp$logFC < 1)))


for (i in sigRes_HTSeq_2_Colnodysp$logFC) {
  # print(i)
  if (i > 1) {
    new_row <- data.frame(regulation = "up")
    deg_reg <- rbind(deg_reg,new_row)
  } else {new_row <- data.frame(regulation = "down")
    deg_reg <- rbind(deg_reg,new_row)}
  }

sigRes_HTSeq_2_Colnodysp_reg <- cbind(sigRes_HTSeq_2_Colnodysp, deg_reg)
nrow(sigRes_HTSeq_2_Colnodysp_reg)
```


## Visualize results in a Volcano plot

```{r}
library(ggplot2)
library(ggrepel)

volcanoPlot <- ggplot(res$table,
                      aes(x = logFC, y = -log10(FDR),
                          color = ifelse(FDR < 0.05 & abs(logFC) > 1, "darkred", "grey"))) +
  geom_point() +
  xlab(expression("Fold Change, Log"[2]*"")) +
  ylab(expression("Adjusted P value, Log"[10]*"")) +
  geom_vline(xintercept = c(-1, 1), linetype = "dotted", linewidth = 1) +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", linewidth = 1) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_colour_manual(values = c("darkred", "grey", "steelblue")) +
  geom_text_repel(aes(x = logFC, y = -log10(FDR), label = rownames(res$table)[1:10],
                      size = 2, color = "steelblue"),
                  data = res$table[1:10, ])
print(volcanoPlot)
```

### 1. Perform hypothesis testing

### Colitisonly vs coldysp

```{r}
(res <- glmQLFTest(fit, contrast = contrastMatrix2Coldysp))
nrow(res)
```

### 2. filter genes, i.e., (i) create a table of top DEG (Tags) sorted by Pvalue

```{r}
(res <- topTags(res, n = nrow(countTable)))
nrow(res)
```

### ... (ii) based on the level of significance (adjusted Pvalue, FDR adjusted Pvalue of 0.05) and the level of relevance (log2FC \> 1, i.e., FC of at least 2)

```{r}
sigRes <- subset(res$table, FDR < 0.05 & abs(logFC) > 1)
sigRes_HTSeq_2_Coldysp <- sigRes
knitr::kable(head(sigRes))
# htseq_tags <- rownames(res$table)[1:20]; htseq_tags
nrow(sigRes)
# rownames(sigRes)[1:20]
```



## To add a column of up- and down-regulated genes

```{r}
deg_reg <- data.frame("regulation" = NULL)

#to confirm the number of DEG is correct
nrow(sigRes_HTSeq_2_Coldysp) ; sum(nrow(subset(sigRes_HTSeq_2_Coldysp,sigRes_HTSeq_2_Coldysp$logFC > 1)),nrow(subset(sigRes_HTSeq_2_Coldysp,sigRes_HTSeq_2_Coldysp$logFC < 1)))


for (i in sigRes_HTSeq_2_Coldysp$logFC) {
  # print(i)
  if (i > 1) {
    new_row <- data.frame(regulation = "up")
    deg_reg <- rbind(deg_reg,new_row)
  } else {new_row <- data.frame(regulation = "down")
    deg_reg <- rbind(deg_reg,new_row)}
  }

sigRes_HTSeq_2_Coldysp_reg <- cbind(sigRes_HTSeq_2_Coldysp, deg_reg)
nrow(sigRes_HTSeq_2_Coldysp_reg)
```


## Visualize results in a Volcano plot

```{r}
library(ggplot2)
library(ggrepel)

volcanoPlot <- ggplot(res$table,
                      aes(x = logFC, y = -log10(FDR),
                          color = ifelse(FDR < 0.05 & abs(logFC) > 1, "darkred", "grey"))) +
  geom_point() +
  xlab(expression("Fold Change, Log"[2]*"")) +
  ylab(expression("Adjusted P value, Log"[10]*"")) +
  geom_vline(xintercept = c(-1, 1), linetype = "dotted", linewidth = 1) +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", linewidth = 1) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_colour_manual(values = c("darkred", "grey", "steelblue")) +
  geom_text_repel(aes(x = logFC, y = -log10(FDR), label = rownames(res$table)[1:10],
                      size = 2, color = "steelblue"),
                  data = res$table[1:10, ])
print(volcanoPlot)
```

### 1. Perform hypothesis testing

### colnodysp vs coldysp

```{r}
(res <- glmQLFTest(fit, contrast = contrastMatrix3Colnodysp))
nrow(res)
```

### 2. filter genes, i.e., (i) create a table of top DEG (Tags) sorted by Pvalue

```{r}
(res <- topTags(res, n = nrow(countTable)))
nrow(res)
```

### ... (ii) based on the level of significance (adjusted Pvalue, FDR adjusted Pvalue of 0.05) and the level of relevance (log2FC \> 1, i.e., FC of at least 2)

```{r}
sigRes <- subset(res$table, FDR < 0.05 & abs(logFC) > 1)
sigRes_HTSeq_3_Colnodysp <- sigRes
knitr::kable(head(sigRes))
# htseq_tags <- rownames(res$table)[1:20]; htseq_tags
nrow(sigRes)
# rownames(sigRes)[1:20]
```



## To add a column of up- and down-regulated genes

```{r}
deg_reg <- data.frame("regulation" = NULL)

#to confirm the number of DEG is correct
nrow(sigRes_HTSeq_3_Colnodysp) ; sum(nrow(subset(sigRes_HTSeq_3_Colnodysp,sigRes_HTSeq_3_Colnodysp$logFC > 1)),nrow(subset(sigRes_HTSeq_3_Colnodysp,sigRes_HTSeq_3_Colnodysp$logFC < 1)))


for (i in sigRes_HTSeq_3_Colnodysp$logFC) {
  # print(i)
  if (i > 1) {
    new_row <- data.frame(regulation = "up")
    deg_reg <- rbind(deg_reg,new_row)
  } else {new_row <- data.frame(regulation = "down")
    deg_reg <- rbind(deg_reg,new_row)}
  }

sigRes_HTSeq_3_Colnodysp_reg <- cbind(sigRes_HTSeq_3_Colnodysp, deg_reg)
```

## Visualize results in a Volcano plot

```{r}
library(ggplot2)
library(ggrepel)

volcanoPlot <- ggplot(res$table,
                      aes(x = logFC, y = -log10(FDR),
                          color = ifelse(FDR < 0.05 & abs(logFC) > 1, "darkred", "grey"))) +
  geom_point() +
  xlab(expression("Fold Change, Log"[2]*"")) +
  ylab(expression("Adjusted P value, Log"[10]*"")) +
  geom_vline(xintercept = c(-1, 1), linetype = "dotted", linewidth = 1) +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", linewidth = 1) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_colour_manual(values = c("darkred", "grey", "steelblue")) +
  geom_text_repel(aes(x = logFC, y = -log10(FDR), label = rownames(res$table)[1:10],
                      size = 2, color = "steelblue"),
                  data = res$table[1:10, ])
print(volcanoPlot)
```


## Save the results as an R image

```{r}
save.image("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/hisat2_htseq_pipeline/deseq2_statistical_analysis_HTSeq.Rdata")
```



=======================================================================================================
# to compare the FDR and Pvalue as thresholds for significance; to help answer the question whether both threshold give consistent results


## Step 4: Perform hypothesis testing
### 1. Perform hypothesis testing
```{r}
res_Col <- glmQLFTest(fit, contrast = contrastMatrixCol)
res_Colnodysp <- glmQLFTest(fit, contrast = contrastMatrixColnodysp)
res_Coldysp <- glmQLFTest(fit, contrast = contrastMatrixColdysp)

```

### 2. filter genes, i.e., (i) create a table of top DEG (Tags) sorted by Pvalue


```{r}
res_Col <- glmQLFTest(fit, contrast = contrastMatrixCol)

res <- res_Col

res <- topTags(res, n = nrow(countTable))
sigRes1 <- subset(res$table, FDR < 0.05 & abs(logFC) > 1)
# knitr::kable(head(sigRes))

sigRes2 <- subset(res$table, PValue < 0.05 & FDR < 0.05 & abs(logFC)> 1)
# knitr::kable(head(sigRes))

sigRes3 <- subset(res$table, PValue < 0.05 & abs(logFC)> 1)

# knitr::kable(table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3)))

table1 <- table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3))
t1 <- paste("noCol vs ColOnly", knitr::kable(table1))
print(t1)

knitr::kable(table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3)))
```


```{r}
res_Colnodysp <- glmQLFTest(fit, contrast = contrastMatrixColnodysp)

res <- res_Colnodysp

res <- topTags(res, n = nrow(countTable))
sigRes1 <- subset(res$table, FDR < 0.05 & abs(logFC) > 1)
# knitr::kable(head(sigRes))

sigRes2 <- subset(res$table, PValue < 0.05 & FDR < 0.05 & abs(logFC)> 1)
# knitr::kable(head(sigRes))

sigRes3 <- subset(res$table, PValue < 0.05 & abs(logFC)> 1)

# knitr::kable(table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3)))

table2 <- table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3))
t2 <- paste("noCol vs Colnodysp", knitr::kable(table2))
print(t2)
knitr::kable(table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3)))

```

### ... (ii) based on the level of significance (adjusted Pvalue, FDR adjusted Pvalue of 0.05) and the level of relevance (log2FC > 1, i.e., FC of at least 2)

```{r}
res_Coldysp <- glmQLFTest(fit, contrast = contrastMatrixColdysp)

res <- res_Coldysp


res <- topTags(res, n = nrow(countTable))
sigRes1 <- subset(res$table, FDR < 0.05 & abs(logFC) > 1)
# knitr::kable(head(sigRes))

sigRes2 <- subset(res$table, PValue < 0.05 & FDR < 0.05 & abs(logFC)> 1)
# knitr::kable(head(sigRes))

sigRes3 <- subset(res$table, PValue < 0.05 & abs(logFC)> 1)

# knitr::kable(table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3)))

table3 <- table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3))
t3 <- paste("noCol vs ColOnly", knitr::kable(table3))
print(t3)

knitr::kable(table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3)))
```


```{r}
sigRes1 <- subset(res$table, FDR < 0.05 & abs(logFC) > 1)
sigRes_HTG_Coldysp <- sigRes
knitr::kable(head(sigRes))
nrow(sigRes1)

sigRes2 <- subset(res$table, PValue < 0.05 & FDR < 0.05 & abs(logFC)> 1)
sigRes_HTG_Coldysp <- sigRes
knitr::kable(head(sigRes))
nrow(sigRes2)

sigRes3 <- subset(res$table, PValue < 0.05 & abs(logFC)> 1)
nrow(sigRes3)

knitr::kable(table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3)))



paste("noCol vs ColOnly", knitr::kable(table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3))))
```




```{r}

# for i in 
# sigRes1 <- subset(res$table, FDR < 0.05 & abs(logFC) > 1)
# sigRes_HTG_Coldysp <- sigRes
# knitr::kable(head(sigRes))
# nrow(sigRes1)
# 
# sigRes2 <- subset(res$table, PValue < 0.05 & FDR < 0.05 & abs(logFC)> 1)
# sigRes_HTG_Coldysp <- sigRes
# knitr::kable(head(sigRes))
# nrow(sigRes2)
# 
# sigRes3 <- subset(res$table, PValue < 0.05 & abs(logFC)> 1)
# nrow(sigRes3)
# 
# knitr::kable(table(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3)))
# 
# 
# 
# paste("noCol vs ColOnly", knitr::kable(table1)(FDR = nrow(sigRes1), FDRnPvalue = nrow(sigRes2), Pvalue = nrow(sigRes3))))
```


