---
title: "To determine the filtering criteria"
author: "Hildah Njoroge"
date: "2023-08-18"
output: html_document
---

# HISAT2-HTSseq data

### Load the data from the preprocessing step

```{r}

load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/hisat2_htseq_pipeline/htseq_dds_filtering.Rdata")


load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/htg_pipeline/htg_dds_filtering.Rdata")

```

### Distribution

##### To determine the threshold to set for filtering low expression values

```{r}
# hist(assay(normCounts_vst), main = "HISAT2-HTSeq Pipeline output when VST normalised",xlim = c(0,20), breaks = seq(0, 20, by = 0.8), ylim = c(0,15000))

meanLog2CPM <- rowMeans(log2(cpm(counts(ddsHTSeq)) + 1))
meanLog2CPM1 <- log2(rowMeans(cpm(counts(ddsHTSeq)) + 1))
meanLog2CPM2 <-  aveLogCPM(y = counts(ddsHTSeq),)

hist(meanLog2CPM)
hist(meanLog2CPM, main = "HISAT2-HTSeq Pipeline output before log2 CPM normalisation",xlim = c(0,20), breaks = seq(0, 20, by = 0.1), ylim = c(0,60))

hist(meanLog2CPM1, main = "HISAT2-HTSeq Pipeline output before log2 CPM normalisation",xlim = c(0,20), breaks = seq(0, 20, by = 0.1), ylim = c(0,60))

hist(meanLog2CPM2, main = "HISAT2-HTSeq Pipeline output before log2 CPM normalisation",xlim = c(-3,20), breaks = seq(-3, 20, by = 0.1), ylim = c(0,60))

hist(meanLog2CPM, main = "HISAT2-HTSeq Pipeline output before log2 CPM normalisation",xlim = c(0,20), breaks = seq(0, 20, by = 0.1), ylim = c(0,60))


# expr_cutoff=0.05
expr_cutoff=0.1
abline(v = expr_cutoff, col = "red", lwd = 1)

# counts(ddsHTSeq[meanLog2CPM > 1,])
# paste(sum(meanLog2CPM <= 0.05), "genes are filtered")
# sum(meanLog2CPM > 0.05)
# cpm_filtered_htseq <- counts(ddsHTSeq[meanLog2CPM > 0.05,])
paste(sum(meanLog2CPM <= 0.1), "genes are filtered")
sum(meanLog2CPM1 > 0.1)
cpm_filtered_htseq <- counts(ddsHTSeq[meanLog2CPM > 0.1,])

cpm_filtered_htseq_meanLog2CPM <- rowMeans(log2(cpm(cpm_filtered_htseq) + 1))

hist(cpm_filtered_htseq_meanLog2CPM)
hist(cpm_filtered_htseq_meanLog2CPM, main = "HISAT2-HTSeq Pipeline output when log2 CPM normalised",xlim = c(0,20), breaks = seq(0, 20, by = 0.1), ylim = c(0,60))



expr_cutoff=0.05
abline(v = expr_cutoff, col = "red", lwd = 1)

```

#### Defined the DDS using the new cpm filtering criteria - removing the left peak, by setting the threshold at 0.05. This threshold removed 39,210 genes. 

### BUT, on 23/10/23, I have adjusted the threshold to 0.1 up from 0.05. 3 genes in the htseq pipeline are excluded and still, non from the HTG pipeline

##### 1669 genes have been retained- This is so as to retain at least as many genes as those of the HTG panel


```{r}
library(RColorBrewer)
library(pheatmap)
#Prepare data for QC after method 2 prefiltering
dds_cpm_filtered_HTSeq1 <- DESeqDataSetFromMatrix(as.matrix(cpm_filtered_htseq),
                              design = ~ condition,
                              colData = sampleTable_shortened)

paste("the number of gene features and samples,respectively are,",dim(dds_cpm_filtered_HTSeq1))
normCounts_vst <- vst(dds_cpm_filtered_HTSeq1, blind = TRUE) ; normCounts_vst_htseq <- normCounts_vst
hist(assay(normCounts_vst), main = "HISAT2-HTSeq Pipeline output when\
     VST normalised after log2(cpm) > 0.05 filtering ", xlim = c(5,20), breaks = seq(0, 20, by = 0.5))


qqnorm(assay(normCounts_vst), main = "Normal Q-Q Plot for the HISAT2-HTSeq count data")
qqline(assay(normCounts_vst))


# sampleTable$condition <- relevel(sampleTable$condition, ref = "noColitis")
sampleColor <- brewer.pal(4, "Accent")[1:4]
# # sampleColor <- cols <- c( "red","orange", "#0080FF","lightgreen")
names(sampleColor) <- levels(sampleTable_shortened$condition)
# sampleTable$condition <- factor(sampleTable$condition)
sampleDist_vst <- cor(assay(normCounts_vst), method = "pearson")

rownames(sampleDist_vst) <- NULL
pheatmap(sampleDist_vst,
         clustering_distance_rows = as.dist(1 - sampleDist_vst),
         clustering_distance_cols = as.dist(1 - sampleDist_vst),
         annotation_col = data.frame(Disease = sampleTable_shortened$condition,
                                     row.names = sampleTable_shortened$sampleName),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap after filtering by log2(cpm) > 0.05 &\
         VST normalised approach - Pearson correlation")




# Heatmap after filtering by log2(cpm) > 0.05 & VST normalised approach - Spearman correlation
sampleDist_vst <- cor(assay(normCounts_vst), method = "spearman")

rownames(sampleDist_vst) <- NULL
pheatmap(sampleDist_vst,
         clustering_distance_rows = as.dist(1 - sampleDist_vst),
         clustering_distance_cols = as.dist(1 - sampleDist_vst),
         annotation_col = data.frame(Disease = sampleTable_shortened$condition,
                                     row.names = sampleTable_shortened$sampleName),
         annotation_colors = list(Disease = sampleColor),
         main = "A sample-sample correlation heatmap of the HISAT2-HTSeq\
         pipeline data by Spearman's correlation")


sampleDist_vst <- cor(assay(normCounts_vst), method = "spearman")

rownames(sampleDist_vst) <- NULL  # to remove the sample names on the column 
pheatmap(sampleDist_vst,
         clustering_distance_rows = as.dist(sampleDist_vst),
         clustering_distance_cols = as.dist(sampleDist_vst),
         annotation_col = data.frame(Disease = sampleTable_shortened$condition,
                                     row.names = sampleTable_shortened$sampleName),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap after filtering by log2(cpm) > 0.1 &\
         VST normalised approach - Spearman correlation")
```


#### on 02.03.2024 --> Attempting to modify the "sampleName" column of the sampleTable file in the HTÌSAT2-HTSeq metadata file, to shorten the names by removing the "_counts.txt"

```{r}
sampleTable_shortened <- sampleTable
sampleTable_shortened$sampleName <- gsub("_counts.txt", "", sampleTable_shortened$sampleName)
sampleTable_shortened$fileName <- gsub("_counts.txt", "", sampleTable_shortened$fileName)
head(sampleTable_shortened)
```

```{r}
library(RColorBrewer)
library(pheatmap)
#Prepare data for QC after method 2 prefiltering
dds_cpm_filtered_HTSeq1 <- DESeqDataSetFromMatrix(as.matrix(cpm_filtered_htseq),
                              design = ~ condition,
                              colData = sampleTable_shortened)

paste("the number of gene features and samples,respectively are,",dim(dds_cpm_filtered_HTSeq1))
normCounts_vst <- vst(dds_cpm_filtered_HTSeq1, blind = TRUE) ; normCounts_vst_htseq <- normCounts_vst
hist(assay(normCounts_vst), main = "HISAT2-HTSeq Pipeline output when\
     VST normalised after log2(cpm) > 0.05 filtering ", xlim = c(5,20), breaks = seq(0, 20, by = 0.5))


qqnorm(assay(normCounts_vst), main = "Normal Q-Q Plot for the HISAT2-HTSeq count data")
qqline(assay(normCounts_vst))


# sampleTable$condition <- relevel(sampleTable$condition, ref = "noColitis")
sampleColor <- brewer.pal(4, "Accent")[1:4]
# # sampleColor <- cols <- c( "red","orange", "#0080FF","lightgreen")
names(sampleColor) <- levels(sampleTable_shortened$condition)
# sampleTable$condition <- factor(sampleTable$condition)
sampleDist_vst <- cor(assay(normCounts_vst), method = "pearson")

rownames(sampleDist_vst) <- NULL
pheatmap(sampleDist_vst,
         clustering_distance_rows = as.dist(1 - sampleDist_vst),
         clustering_distance_cols = as.dist(1 - sampleDist_vst),
         annotation_col = data.frame(Disease = sampleTable_shortened$condition,
                                     row.names = sampleTable_shortened$sampleName),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap after filtering by log2(cpm) > 0.05 &\
         VST normalised approach - Pearson correlation")




# Heatmap after filtering by log2(cpm) > 0.05 & VST normalised approach - Spearman correlation
sampleDist_vst <- cor(assay(normCounts_vst), method = "spearman")

rownames(sampleDist_vst) <- NULL
pheatmap(sampleDist_vst,
         clustering_distance_rows = as.dist(1 - sampleDist_vst),
         clustering_distance_cols = as.dist(1 - sampleDist_vst),
         annotation_col = data.frame(Disease = sampleTable_shortened$condition,
                                     row.names = sampleTable_shortened$sampleName),
         annotation_colors = list(Disease = sampleColor),
         main = "A sample-sample correlation heatmap of the HISAT2-HTSeq\
         pipeline data by Spearman's correlation")


sampleDist_vst <- cor(assay(normCounts_vst), method = "spearman")

rownames(sampleDist_vst) <- NULL  # to remove the sample names on the column 
pheatmap(sampleDist_vst,
         clustering_distance_rows = as.dist(sampleDist_vst),
         clustering_distance_cols = as.dist(sampleDist_vst),
         annotation_col = data.frame(Disease = sampleTable_shortened$condition,
                                     row.names = sampleTable_shortened$sampleName),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap after filtering by log2(cpm) > 0.1 &\
         VST normalised approach - Spearman correlation")
```

```{r}
library(ggplot2)
library(ggrepel)
library(RColorBrewer)
par(mfrow=c(3,1))
# pcaRes <- prcomp(t(assay(normCounts_rlog)))
# varExp <- round(pcaRes$sdev^2 / sum(pcaRes$sdev^2) * 100)
# pcaDF <- data.frame(
#   PC1 = pcaRes$x[, 1],
#   PC2 = pcaRes$x[, 2],
#   Disease = coldata$condition2,
#   Sample = row.names(coldata))
# pcaPlot <- ggplot(
#   data = pcaDF,
#   mapping = aes(x = PC1, y = PC2, color = Disease, label = Sample)) +
#   geom_point(size = 3) +
#   geom_text_repel(size = 0) +
#   labs(x = paste0("PC1 (", varExp[1], " %)"), y = paste0("PC2 (", varExp[2], " %)")) +
#   theme_minimal() +
#   theme(axis.text = element_text(size = 12), legend.text = element_text(size = 10)) +
#   scale_color_manual(values = brewer.pal(4, "Accent")[1:4])
# print(pcaPlot)

pcaRes <- prcomp(t(assay(normCounts_vst_htseq)))
varExp <- round(pcaRes$sdev^2 / sum(pcaRes$sdev^2) * 100)
pcaDF <- data.frame(
  PC1 = pcaRes$x[, 1],
  PC2 = pcaRes$x[, 2],
  Disease = coldata$condition2,
  Sample = row.names(coldata))
pcaPlot <- ggplot(
  data = pcaDF,
  mapping = aes(x = PC1, y = PC2, color = Disease, label = Sample)) +
  geom_point(size = 3) +
  geom_text_repel(size = 0) +
  labs(x = paste0("PC1 (", varExp[1], " %)"), y = paste0("PC2 (", varExp[2], " %)")) +
  theme_minimal() +
  theme(axis.text = element_text(size = 12), legend.text = element_text(size = 10)) +
  scale_color_manual(values = brewer.pal(4, "Accent")[1:4])+
  ggtitle("PCA Plot of the VST-transformed HISAT2-HTSeq count data")
print(pcaPlot)




pcaPlot <- ggplot(
  data = pcaDF,
  mapping = aes(x = PC1, y = PC2, color = Disease, label = Sample)) +
  geom_point(size = 3) +
  geom_text_repel(size = 3) +
  labs(x = paste0("PC1 (", varExp[1], " %)"), y = paste0("PC2 (", varExp[2], " %)")) +
  theme_minimal() +
  theme(axis.text = element_text(size = 12), legend.text = element_text(size = 10)) +
  scale_color_manual(values = brewer.pal(4, "Accent")[1:4])+
  ggtitle("PCA Plot of the VST-transformed HISAT2-HTSeq count data")
print(pcaPlot)




# to run PCA on disease group only
pcaRes <- prcomp(t(assay(normCounts_vst_htseq)[,1:42])) 
varExp <- round(pcaRes$sdev^2 / sum(pcaRes$sdev^2) * 100)
pcaDF <- data.frame(
  PC1 = pcaRes$x[, 1],
  PC2 = pcaRes$x[, 2],
  Disease = coldata[1:42,]$condition2,
  Sample = row.names(coldata[1:42,]))

pcaPlot <- ggplot(
  data = pcaDF,
  mapping = aes(x = PC1, y = PC2, color = Disease, label = Sample)) +
  geom_point(size = 3) +
  geom_text_repel(size = 3) +
  labs(x = paste0("PC1 (", varExp[1], " %)"), y = paste0("PC2 (", varExp[2], " %)")) +
  theme_minimal() +
  theme(axis.text = element_text(size = 12), legend.text = element_text(size = 10)) +
  scale_color_manual(values = brewer.pal(4, "Accent")[1:4])+
  ggtitle("PCA Plot of the VST-transformed HISAT2-HTSeq count data")
print(pcaPlot)

```

=================================================================================================

# HTG data

### Load the data from the preprocessing step

```{r}

load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/htg_pipeline/htg_dds_filtering.Rdata")

```

### Distribution

##### To determine the threshold to set for filtering low expression values

```{r}
# hist(assay(normCounts_vst), main = "HISAT2-HTSeq Pipeline output when VST normalised",xlim = c(0,20), breaks = seq(0, 20, by = 0.8), ylim = c(0,15000))

meanLog2CPM <- rowMeans(log2(cpm(counts(ddsHTG)) + 1))
hist(meanLog2CPM)
hist(meanLog2CPM, main = "HTG Pipeline output before log2 CPM normalisation",xlim = c(0,20), breaks = seq(0, 20, by = 0.1), ylim = c(0,60))

hist(meanLog2CPM, main = "HTG Pipeline output before log2 CPM normalisation",xlim = c(0,20), breaks = seq(0, 20, by = 0.1), ylim = c(0,60))


expr_cutoff=0.1
abline(v = expr_cutoff, col = "red", lwd = 1)

# counts(ddsHTG[meanLog2CPM > 1,])
# paste(sum(meanLog2CPM <= 0.05), "genes are filtered")
# sum(meanLog2CPM > 0.05)
# cpm_filtered_htg <- counts(ddsHTG[meanLog2CPM > 0.05,])

paste(sum(meanLog2CPM <= 0.1), "genes are filtered")
sum(meanLog2CPM > 0.1)
cpm_filtered_htg <- counts(ddsHTG[meanLog2CPM > 0.1,])

cpm_filtered_htg_meanLog2CPM <- rowMeans(log2(cpm(cpm_filtered_htg) + 1))

hist(cpm_filtered_htg_meanLog2CPM)
hist(cpm_filtered_htg_meanLog2CPM, main = "HTG Pipeline output when log2 CPM normalised",xlim = c(0,20), breaks = seq(0, 20, by = 0.1), ylim = c(0,40))

expr_cutoff=0.1
abline(v = expr_cutoff, col = "red", lwd = 1)

```

#### Defined the DDS using the new cpm filtering criteria - removing the left peak, by setting the threshold at 0.05. This threshold removed 39,210 genes.

##### 1669 genes have been retained- This is so as to retain at least as many genes as those of the HTG panel

### BUT, after adjusting the threshold to 0.1 meanLogCPM up from 0.05, 1666 genes are retained -- this was done on the 23/10/23

```{r}
library(RColorBrewer)
library(pheatmap)
#Prepare data for QC after method 2 prefiltering
dds_cpm_filtered_HTG1 <- DESeqDataSetFromMatrix(as.matrix(cpm_filtered_htg),
                              design = ~ condition2,
                              colData = coldata)

paste("the number of gene features and samples,respectively are,",dim(dds_cpm_filtered_HTG1))
normCounts_vst <- vst(dds_cpm_filtered_HTG1, blind = TRUE) ; normCounts_vst_htg <- normCounts_vst
hist(assay(normCounts_vst), main = "HTG Pipeline output when\
     VST normalised after log2(cpm) > 0.05 filtering ", xlim = c(0,20), breaks = seq(0, 20, by = 0.8))

qqnorm(assay(normCounts_vst), main = "Normal Q-Q Plot for the HTG count data")
qqline(assay(normCounts_vst))

# sampleTable$condition <- relevel(sampleTable$condition, ref = "noColitis")
sampleColor <- brewer.pal(4, "Accent")[1:4]
# # sampleColor <- cols <- c( "red","orange", "#0080FF","lightgreen")
# names(sampleColor) <- levels(sampleTable$condition)
names(sampleColor) <- levels(coldata$condition2)
# sampleTable$condition <- factor(sampleTable$condition)
sampleDist_vst <- cor(assay(normCounts_vst), method = "pearson")

rownames(sampleDist_vst) <- NULL
pheatmap(sampleDist_vst,
         clustering_distance_rows = as.dist(1 - sampleDist_vst),
         clustering_distance_cols = as.dist(1 - sampleDist_vst),
         annotation_col = data.frame(Disease = coldata$condition2,
                                     row.names = row.names(coldata)),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap after filtering by log2(cpm) > 0.05 &\
         VST normalised approach - Pearson correlation")



# Heatmap after filtering by log2(cpm) > 0.05 & VST normalised approach - Spearman correlation
sampleDist_vst <- cor(assay(normCounts_vst), method = "spearman")
rownames(sampleDist_vst) <- NULL
pheatmap(sampleDist_vst,
         clustering_distance_rows = as.dist(1 - sampleDist_vst),
         clustering_distance_cols = as.dist(1 - sampleDist_vst),
         annotation_col = data.frame(Disease = coldata$condition2,
                                     row.names = row.names(coldata)),
         annotation_colors = list(Disease = sampleColor),
         main = "A sample-sample correlation heatmap of the HTG \
         pipeline data by Spearman's correlation")
```

```{r}
library(ggplot2)
library(ggrepel)
library(RColorBrewer)
par(mfrow=c(3,1))
# pcaRes <- prcomp(t(assay(normCounts_rlog)))
# varExp <- round(pcaRes$sdev^2 / sum(pcaRes$sdev^2) * 100)
# pcaDF <- data.frame(
#   PC1 = pcaRes$x[, 1],
#   PC2 = pcaRes$x[, 2],
#   Disease = coldata$condition2,
#   Sample = row.names(coldata))
# pcaPlot <- ggplot(
#   data = pcaDF,
#   mapping = aes(x = PC1, y = PC2, color = Disease, label = Sample)) +
#   geom_point(size = 3) +
#   geom_text_repel(size = 0) +
#   labs(x = paste0("PC1 (", varExp[1], " %)"), y = paste0("PC2 (", varExp[2], " %)")) +
#   theme_minimal() +
#   theme(axis.text = element_text(size = 12), legend.text = element_text(size = 10)) +
#   scale_color_manual(values = brewer.pal(4, "Accent")[1:4])
# print(pcaPlot)

pcaRes <- prcomp(t(assay(normCounts_vst)))
varExp <- round(pcaRes$sdev^2 / sum(pcaRes$sdev^2) * 100)
pcaDF <- data.frame(
  PC1 = pcaRes$x[, 1],
  PC2 = pcaRes$x[, 2],
  Disease = coldata$condition2,
  Sample = row.names(coldata))
pcaPlot <- ggplot(
  data = pcaDF,
  mapping = aes(x = PC1, y = PC2, color = Disease, label = Sample)) +
  geom_point(size = 3) +
  geom_text_repel(size = 0) +
  labs(x = paste0("PC1 (", varExp[1], " %)"), y = paste0("PC2 (", varExp[2], " %)")) +
  theme_minimal() +
  theme(axis.text = element_text(size = 12), legend.text = element_text(size = 10)) +
  scale_color_manual(values = brewer.pal(4, "Accent")[1:4]) +
  ggtitle("PCA Plot of the VST-transformed HTG count data")
print(pcaPlot)






# to run PCA on disease group only
pcaRes <- prcomp(t(assay(normCounts_vst)[,1:42])) 
varExp <- round(pcaRes$sdev^2 / sum(pcaRes$sdev^2) * 100)
pcaDF <- data.frame(
  PC1 = pcaRes$x[, 1],
  PC2 = pcaRes$x[, 2],
  Disease = coldata[1:42,]$condition2,
  Sample = row.names(coldata[1:42,]))

pcaPlot <- ggplot(
  data = pcaDF,
  mapping = aes(x = PC1, y = PC2, color = Disease, label = Sample)) +
  geom_point(size = 3) +
  geom_text_repel(size = 1) +
  labs(x = paste0("PC1 (", varExp[1], " %)"), y = paste0("PC2 (", varExp[2], " %)")) +
  theme_minimal() +
  theme(axis.text = element_text(size = 12), legend.text = element_text(size = 10)) +
  scale_color_manual(values = brewer.pal(4, "Accent")[1:4]) +
  ggtitle("PCA Plot of the VST-transformed HTG count data")
print(pcaPlot)
```

### save the data as an R image

```{r}
save.image(file = "C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/filtering_criteria.Rdata")
```

=================================================================================== \### Exploring heatmap plotting on the HTG data

```{r}
normCounts_vst <- vst(dds_cpm_filtered_HTG1, blind = TRUE)
sampleDist_vst <- cor(assay(normCounts_vst), method = "spearman")
rownames(sampleDist_vst) <- NULL
pheatmap(sampleDist_vst,
         clustering_distance_rows = as.dist(1-sampleDist_vst),
         clustering_distance_cols = as.dist(1-sampleDist_vst),
         annotation_col = data.frame(Disease = coldata$condition2,
                                     row.names = row.names(coldata)),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap after filtering by log2(cpm) > 0.05 &\
         VST normalised approach - Spearman correlation")
```

### using the "corrplot" to plot the correlation matrix

```{r}
sampleDist_vst <- cor(assay(normCounts_vst), method = "spearman")
corrplot(sampleDist_vst)
corrplot(sampleDist_vst,order = "hclust")
corrplot(sampleDist_vst, is.corr = F)
corrplot(sampleDist_vst, is.corr = F, order = "hclust")
# corrplot(sampleDist_vst,order = "hclust",method = "number")
```

## Euclidean distance calculation

#### by first having calculated the Euclidean Distance then plottin the heatmap

```{r}
euclidDist_vst <- dist(t(assay(normCounts_vst)), method = "euclidean")
euclidDist_vst <- as.matrix(euclidDist_vst)
corrplot(euclidDist_vst,is.corr = F)

rownames(euclidDist_vst) <- NULL
pheatmap(euclidDist_vst,
         clustering_distance_cols = as.dist(euclidDist_vst),
         clustering_distance_rows = as.dist(euclidDist_vst),
         annotation_col = data.frame(Disease = coldata$condition2,
                                     row.names = row.names(coldata)),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap after filtering by log2(cpm) > 0.05 &\
         VST normalised approach - Euclidean Dist")
```

#### in this plot, I have just run 'pheatmap' giving it the VST-normlised HTG counts as the input. the default method for calculating hte distances is 'Euclidean' and complete linkage of defining the hierarchical clusters.

```{r}
normCounts_vst <- vst(dds_cpm_filtered_HTG1, blind = TRUE)
trial <- assay(normCounts_vst)
dim(trial)
rownames(trial) <- NULL
# colnames(trial) <- paste((sub("sample", "S", colnames(assay(normCounts_vst)))))
## Heatmap after filtering by log2(cpm) > 0.05 & VST normalisation using pheatmap default parameters

pheatmap(trial,
         annotation_col = data.frame(Disease = coldata$condition2,
                                     row.names = row.names(coldata)),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap with Hierarchical Clustering of Gene Features (y-axis) \
         and Sample (x-axis) Similarity - Euclidean Distances ")


pheatmap(t(trial),
         annotation_row = data.frame(Disease = coldata$condition2,
                                     row.names = row.names(coldata)),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap with Hierarchical Clustering of Sample (y-axis) \
         and Gene Features (x-axis) Similarity - Euclidean Distances ")
```

=================================================================================== \### Exploring heatmap plotting on the HTSeq data

```{r}
normCounts_vst <- vst(dds_cpm_filtered_HTSeq1, blind = TRUE)
sampleDist_vst <- cor(assay(normCounts_vst), method = "spearman")
rownames(sampleDist_vst) <- NULL
pheatmap(sampleDist_vst,
         clustering_distance_rows = as.dist(1-sampleDist_vst),
         clustering_distance_cols = as.dist(1-sampleDist_vst),
         annotation_col = data.frame(Disease = sampleTable$condition,
                                     row.names = sampleTable$sampleName),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap after filtering by log2(cpm) > 0.05 &\
         VST normalised approach - Spearman correlation")

# https://dputhier.github.io/ASG/practicals/distances/distances.html#hierarchical_clustering_with_hclust
# not sure which is the method to convert the Spearman's correlation to distances
# these two plots give the same result

pheatmap(sampleDist_vst,
         clustering_distance_rows = as.dist((1-sampleDist_vst)/2),
         clustering_distance_cols = as.dist((1-sampleDist_vst)/2),
         annotation_col = data.frame(Disease = sampleTable$condition,
                                     row.names = sampleTable$sampleName),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap after filtering by log2(cpm) > 0.05 &\
         VST normalised approach - Spearman correlation")
```

### using the "corrplot" to plot the correlation matrix

```{r}
sampleDist_vst <- cor(assay(normCounts_vst), method = "spearman")
corrplot(sampleDist_vst)
corrplot(sampleDist_vst,order = "hclust")
corrplot(sampleDist_vst, is.corr = F)
corrplot(sampleDist_vst, is.corr = F, order = "hclust")
# corrplot(sampleDist_vst,order = "hclust",method = "number")
```

## Euclidean distance calculation

#### by first having calculated the Euclidean Distance then plottin the heatmap

```{r}
euclidDist_vst <- dist(t(assay(normCounts_vst)), method = "euclidean")
euclidDist_vst <- as.matrix(euclidDist_vst)
corrplot(euclidDist_vst,is.corr = F)

rownames(euclidDist_vst) <- NULL
pheatmap(euclidDist_vst,
         clustering_distance_cols = as.dist(euclidDist_vst),
         clustering_distance_rows = as.dist(euclidDist_vst),
         annotation_col = data.frame(Disease = sampleTable$condition,
                                     row.names = sampleTable$sampleName),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap after filtering by log2(cpm) > 0.05 &\
         VST normalised approach - Euclidean Dist")
```

#### in this plot, I have just run 'pheatmap' giving it the VST-normlised HTG counts as the input. the default method for calculating hte distances is 'Euclidean' and complete linkage of defining the hierarchical clusters.

```{r}
normCounts_vst <- vst(dds_cpm_filtered_HTSeq1, blind = TRUE)
trial <- assay(normCounts_vst)
dim(trial)
rownames(trial) <- NULL
# colnames(trial) <- paste((sub("sample", "S", colnames(assay(normCounts_vst)))))
## Heatmap after filtering by log2(cpm) > 0.05 & VST normalisation using pheatmap default parameters

pheatmap(trial,
         annotation_col = data.frame(Disease = sampleTable$condition,
                                     row.names = sampleTable$sampleName),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap with Hierarchical Clustering of Gene Features (y-axis) \
         and Sample (x-axis) Similarity - Euclidean Distances ")


pheatmap(t(trial),
         annotation_row = data.frame(Disease = sampleTable$condition,
                                     row.names = sampleTable$sampleName),
         annotation_colors = list(Disease = sampleColor),
         main = "Heatmap with Hierarchical Clustering of Sample (y-axis) \
         and Gene Features (x-axis) Similarity - Euclidean Distances ")
```

=================================================================================== \#### Now, try plotting a correlation plot

### using the "corrplot" to plot the correlation matrix

```{r}
# sampleDist_vst <- cor(assay(normCounts_vst), method = "spearman")
sampleDist_vst <- cor(t(assay(normCounts_vst)), method = "spearman")
corrplot(sampleDist_vst)
corrplot(sampleDist_vst,order = "hclust")
corrplot(sampleDist_vst, is.corr = F)
corrplot(sampleDist_vst, is.corr = F, order = "hclust")
# corrplot(sampleDist_vst,order = "hclust",method = "number")
```

```{r}

sampleDist_vst_sample <- cor(assay(normCounts_vst), method = "spearman")
sampleDist_vst_gene <- cor(t(assay(normCounts_vst)), method = "spearman")
row.names(sampleDist_vst_gene) <- NULL
colnames(sampleDist_vst_gene) <- NULL
pheatmap(sampleDist_vst_gene,
         clustering_distance_cols = as.dist(1 - sampleDist_vst_gene),
         clustering_distance_rows = as.dist(1 - sampleDist_vst_gene),
         main = "Heatmap after filtering by log2(cpm) > 0.05 &\
         VST normalised approach - Spearman's Corr")

pheatmap(sampleDist_vst_gene,
         clustering_distance_cols = as.dist((1-sampleDist_vst_gene)/2),
         clustering_distance_rows = as.dist((1-sampleDist_vst_gene)/2),
         main = "Heatmap after filtering by log2(cpm) > 0.05 &\
         VST normalised approach - Spearman's Corr")

```

```{r}
pear.dist <- as.dist((1 - sampleDist_vst_gene)/2)
hp <- hclust(pear.dist, method="average")
plot(hp, hang=-1, cex=0.2)

pear.dist <- as.dist(1 - sampleDist_vst_gene)
hp <- hclust(pear.dist, method="average")
plot(hp, hang=-1, cex=0.2)

pear.dist <- as.dist((1 - sampleDist_vst_gene)/2)
hp <- hclust(pear.dist, method="complete")
plot(hp, hang=-1, cex=0.2)
```
