---
title: "HTG - HISAT2-HTSeq Raw Count Comparisons"
author: "Hildah Njoroge"
date: "2023-07-28"
output: html_document
---


# HTSeq 
# 1. zero and low expression values filtered HTSeq 

```{r}
load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/hisat2_htseq_pipeline/htseq_dds_filtering.Rdata")

paste("The HISAT2-HTSeq DESeqDataSet")
ddsHTSeq

paste("The cpm_filtered HISAT2-HTSeq DESeqDataSet")
dds_cpm_filtered_HTSeq 

paste("The rowSums_greaterthan_10 HISAT2-HTSeq DESeqDataSet")
dds_rowSums_greaterthan_10_HTSeq


```

# HTG
### versus

```{r}
load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/htg_pipeline/htg_dds_filtering.Rdata")
paste("The HTG DESeqDataSet")
ddsHTG

paste("The cpm_filtered HTG DESeqDataSet")
dds_cpm_filtered_HTG 


paste("The rowSums_greaterthan_10 HTG DESeqDataSet")
dds_rowSums_greaterthan_10_HTG 
```

```{r}
dimensions_of_countTable_HTG <- data.frame(HTG_raw_counts = dim(ddsHTG),
                                       HTG_counts_only_greaterthan10 = dim(dds_rowSums_greaterthan_10_HTG),
                                       HTG_log2_cpm_filtered = dim(dds_cpm_filtered_HTG)) ; knitr::kable(dimensions_of_countTable_HTG)
```

```{r}
dimensions_of_countTable_HTSeq <- data.frame(HTSeq_raw_counts = dim(ddsHTSeq),
                                           HTSeq_counts_only_greaterthan10 = dim(dds_rowSums_greaterthan_10_HTSeq),
                                           HTSeq_log2_cpm_filtered = dim(dds_cpm_filtered_HTSeq)) ; knitr::kable(dimensions_of_countTable_HTSeq)
```

# 2. library size following HISAT2 alignment vs HTG data parser software

# Calculate library size for each sample in the HTG counts

## before normalisation

```{r}
library_sizes_HTGraw <- colSums(counts(ddsHTG))
```

## after log2(cpm) less than 1 filtering

```{r}
library_sizes_HTGcpm <- colSums(counts(dds_cpm_filtered_HTG))
```

## after greater-than-10 filtering --\> similar to raw counts beacause none were filtered

```{r}
library_sizes_HTGgreaterthan10counts <- colSums(counts(dds_rowSums_greaterthan_10_HTG))
```

### versus

# Calculate library size for each sample in the HISAT2-HTSeq counts

## before normalisation

```{r}
library_sizes_HTSEQraw <- colSums(counts(ddsHTSeq))
```

## after log2(cpm) less than 1 filtering

```{r}
library_sizes_HTSEQcpm <- colSums(counts(dds_cpm_filtered_HTSeq))
```

## after greater-than-10 filtering --\> similar to raw counts beacause none were filtered

```{r}
library_sizes_HTSEQgreaterthan10counts <- colSums(counts(dds_rowSums_greaterthan_10_HTSeq))
```

### plot the output

```{r}
boxplot(library_sizes_HTGraw,library_sizes_HTSEQraw,library_sizes_HTGgreaterthan10counts,library_sizes_HTSEQgreaterthan10counts,library_sizes_HTGcpm,library_sizes_HTSEQcpm)
```

## Define the datasets and their labels

## Create a data frame with the datasets

```{r}
datasets <- data.frame(
  HTGraw = library_sizes_HTGraw,
  HTSEQraw = library_sizes_HTSEQraw,
  HTGgreaterthan10counts = library_sizes_HTGgreaterthan10counts,
  HTSEQgreaterthan10counts = library_sizes_HTSEQgreaterthan10counts,
  HTGcpm = library_sizes_HTGcpm,
  HTSEQcpm = library_sizes_HTSEQcpm
)

# Define the datasets and their labels
# datasets_as_list <- list(
#   library_sizes_HTGraw,
#   library_sizes_HTSEQraw,
#   library_sizes_HTGgreaterthan10counts,
#   library_sizes_HTSEQgreaterthan10counts,
#   library_sizes_HTGcpm,
#   library_sizes_HTSEQcpm
# )
# Define the labels
# labels <- c(
#   "HTG Raw",
#   "HTSeq Raw",
#   "HTG >10 Counts",
#   "HTSeq >10 Counts",
#   "HTG CPM",
#   "HTSeq CPM"
# )

labels <- c(
  "HTG",
  "HTSeq",
  "HTG",
  "HTSeq",
  "HTG",
  "HTSeq"
)

```

## Define the colors for the boxes

```{r}
colors <- c("red", "#FF5500", "darkorange", "orange", "green", "lightgreen")
boxplot(datasets, names = labels, col = colors)
labels <- c(
  "HTG Raw",
  "HTSeq Raw",
  "HTG >10 Counts",
  "HTSeq >10 Counts",
  "HTG CPM",
  "HTSeq CPM"
)
```

# Create a legend for the box plot with labels for the plots
## Define the labels
```{r}
boxplot(datasets, names = labels, col = colors)
legend("right", legend = labels, fill = colors)
```
