---
title: "Venn diagram"
author: "Hildah Njoroge"
date: "2023-07-26"
output: html_document
---

# Create a demo data

```{r}
# set.seed(20190708)
# genes <- paste("gene",1:1000,sep="")
# x <- list(
#   A = sample(genes,300), 
#   B = sample(genes,525), 
#   C = sample(genes,440),
#   D = sample(genes,350)
#   )
# 
# head(x$A) ; head(x$B) ; head(x$C) ; head(x$D)
```

# Create Venn diagrams using the ggVennDiagram R package

## Install and load the ggVennDiagram package

### Install

```{r}
if (!require(devtools)) {install.packages("devtools")}
# devtools::install_github("gaospecial/ggVennDiagram") # install from GitHub
```

### Install necessary packages

```{r}

if (!require(devtools)) {install.packages("Rtools")}
if (!require(devtools)) {install.packages("ggVennDiagram")}
```

## Load package

```{r}
library("ggVennDiagram")
```

```{r}
# # Default plot
# ggVennDiagram(x)
# ggVennDiagram(x[1:3])
```

# Now trying with my data

## Load data

```{r}
load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/htg_pipeline/deseq2_statistical_analysis_HTG.Rdata")
```

## Load data

```{r}
load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/hisat2_htseq_pipeline/deseq2_statistical_analysis_HTSeq.Rdata")
```

## create the list

## This represents genes called differentially expressed between pipelines for the colitisOnly group

### Here is a list of the genes that are both significant and relevant as defined by the logFC and FDR

```{r}
genelist <- list(htg = row.names(sigRes_HTG),
          htseq = row.names(sigRes_HTSeq)) ; head(genelist$htg) ; head(genelist$htseq)

genelist_colonly <- genelist 

# # checking for overlap in the top 10 genes called significant and relevant
# genelist <- list(htg = row.names(sigRes_HTG)[1:10],
#           htseq = row.names(sigRes_HTSeq)[1:10]) ; head(genelist$htg) ; head(genelist$htseq)
```

## Plot the Venn D

```{r}
ggVennDiagram(genelist)
ggVennDiagram(genelist, label_alpha = )
ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"))+
  ggplot2::scale_fill_gradient(low="darkred", high = "steelblue")

ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"),show_intersect = TRUE)



```


### to convert the Venn diagram above to a Euler diagram

```{r}
#Load the package
library(VennDiagram)

# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
draw.pairwise.venn(area1=length(genelist$htg), 
                   area2=length(genelist$htseq),
                   cross.area=407, 
                   category=c("HTG pipeline genes", "HTSeq pipeline genes"),
                   fill=c("red","steelblue"),
                   fontface = "bold",
                   cex = 1.8,
                   ext.text = F,
                   cat.just = rep(list(c(0.5,0.4)), 2),
                   cat.dist = rep(0.03, 2),
                   cat.fontface = rep("bold", 2),
                   cat.cex = 1.8,
                   cat.pos = c(-30,30)
)

```

#### An attempt to plot the Venn diagram using Pythom
```{python}
# #instal matplotlib-venn module
# # pip install matplotlib-venn 
# # 
# # import modules
# from matplotlib_venn import venn2
# from matplotlib import pyplot as plt
# 
# # depict venn diagram
# 
# # venn2(subsets = (32,21,407), set_labels = ('Group A', 'Group B'))
# venn2(subsets = (50, 10, 7), set_labels = ('Group A', 'Group B'))
# plt.show()

 
```



## create the list

## This represents genes called differentially expressed between pipelines for the colitis_nodysplasia group

```{r}
genelist <- list(htg = row.names(sigRes_HTG_Colnodysp),
          htseq = row.names(sigRes_HTSeq_Colnodysp)) ; head(genelist$htg) ; head(genelist$htseq)

genelist_colnodys <- genelist
```

## Plot the Venn Diagram

```{r}
ggVennDiagram(genelist)
ggVennDiagram(genelist, label_alpha = )
ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"))+
  ggplot2::scale_fill_gradient(low="darkred", high = "steelblue")

ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"),show_intersect = TRUE)

```

### to convert the Venn diagram above to a Euler diagram

```{r}
# load Venn diagram package 
library("VennDiagram") 

# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
draw.pairwise.venn(area1=length(genelist$htg), 
                   area2=length(genelist$htseq),
                   cross.area=424, 
                   category=c("HTG pipeline genes", "HTSeq pipeline genes"),
                   fill=c("red","steelblue"),
                   fontface = "bold",
                   cex = 1.8,
                   ext.text = F,
                   cat.just = rep(list(c(0.5,0.4)), 2),
                   cat.dist = rep(0.03, 2),
                   cat.fontface = rep("bold", 2),
                   cat.cex = 1.8,
                   cat.pos = c(-30,30)
)

```



## create the list

## This represents genes called differentially expressed between pipelines for the noColitis vs colitis_dysplasia groups

```{r}
genelist <- list(htg = row.names(sigRes_HTG_Coldysp),
          htseq = row.names(sigRes_HTSeq_Coldysp)) ; head(genelist$htg) ; head(genelist$htseq)

genelist_coldys <- genelist

# # checking for overlap in the top 20 genes called significant and relevant
# genelist <- list(htg = row.names(sigRes_HTG_Coldysp)[1:20],
#           htseq = row.names(sigRes_HTSeq_Coldysp)[1:20]) ; head(genelist$htg) head(genelist$htseq)
```

## Plot the Venn Diagram

```{r}
ggVennDiagram(genelist)
ggVennDiagram(genelist, label_alpha = )
ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"))+
  ggplot2::scale_fill_gradient(low="darkred", high = "steelblue")

ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"),show_intersect = TRUE)

```


### to convert the Venn diagram above to a Euler diagram
```{r}
# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
draw.pairwise.venn(area1=length(genelist$htg), 
                   area2=length(genelist$htseq),
                   cross.area=446, 
                   category=c("HTG pipeline genes", "HTSeq pipeline genes"),
                   fill=c("red","steelblue"),
                   fontface = "bold",
                   cex = 1.8,
                   ext.text = F,
                   cat.just = rep(list(c(0.5,0.4)), 2),
                   cat.dist = rep(0.03, 2),
                   cat.fontface = rep("bold", 2),
                   cat.cex = 1.8,
                   cat.pos = c(-30,30)
)

```


## In conclusion, the top number of DEG in the experimental groups compared to the control is the same in both pipelines
## These results are based on the univariate analysis , which involves identifying single biomarkers that are differentially regulated between the groups compared in the analysis


========================================================================================
### contrasts with the reference group being the colitis only group


## create the list

## This represents genes called differentially expressed between pipelines for the Colitisonly vs colitis_nodysplasia groups

```{r}
genelist <- list(htg = row.names(sigRes_HTG_2_Colnodysp),
          htseq = row.names(sigRes_HTSeq_2_Colnodysp)) ; head(genelist$htg) ; head(genelist$htseq)

genelist_2colnodys <- genelist

# # checking for overlap in the top 20 genes called significant and relevant
# genelist <- list(htg = row.names(sigRes_HTG_Coldysp)[1:20],
#           htseq = row.names(sigRes_HTSeq_Coldysp)[1:20]) ; head(genelist$htg) head(genelist$htseq)
```

## Plot the Venn Diagram

```{r}
ggVennDiagram(genelist)
ggVennDiagram(genelist, label_alpha = )
ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"))+
  ggplot2::scale_fill_gradient(low="darkred", high = "steelblue")

ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"),show_intersect = TRUE)

```


### to convert the Venn diagram above to a Euler diagram
```{r}
# load Venn diagram package 
library("VennDiagram") 

# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
draw.pairwise.venn(area1=length(genelist$htg), 
                   area2=length(genelist$htseq),
                   cross.area=25, 
                   category=c("HTG pipeline genes", "HTSeq pipeline genes"),
                   fill=c("red","steelblue"),
                   fontface = "bold",
                   cex = 1.8,
                   ext.text = F,
                   cat.just = rep(list(c(0.5,0.4)), 2),
                   cat.dist = rep(0.03, 2),
                   cat.fontface = rep("bold", 2),
                   cat.cex = 1.8,
                   cat.pos = c(-150,150)
)
```


## create the list

## This represents genes called differentially expressed between pipelines for the Colitisonly vs colitis_dysplasia groups

```{r}
genelist <- list(htg = row.names(sigRes_HTG_2_Coldysp),
          htseq = row.names(sigRes_HTSeq_2_Coldysp)) ; head(genelist$htg) ; head(genelist$htseq)

genelist_2coldys <- genelist

# # checking for overlap in the top 20 genes called significant and relevant
# genelist <- list(htg = row.names(sigRes_HTG_Coldysp)[1:20],
#           htseq = row.names(sigRes_HTSeq_Coldysp)[1:20]) ; head(genelist$htg) head(genelist$htseq)
```

## Plot the Venn Diagram

```{r}
ggVennDiagram(genelist)
ggVennDiagram(genelist, label_alpha = )
ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"))+
  ggplot2::scale_fill_gradient(low="darkred", high = "steelblue")

ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"),show_intersect = TRUE)

```


### to convert the Venn diagram above to a Euler diagram
```{r}
# load Venn diagram package 
library("VennDiagram") 

# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
draw.pairwise.venn(area1=length(genelist$htg), 
                   area2=length(genelist$htseq),
                   cross.area=15, 
                   category=c("HTG pipeline genes", "HTSeq pipeline genes"),
                   fill=c("red","steelblue"),
                   fontface = "bold",
                   cex = 1.8,
                   ext.text = F,
                   cat.just = rep(list(c(0.5,0.4)), 2),
                   cat.dist = rep(0.03, 2),
                   cat.fontface = rep("bold", 2),
                   cat.cex = 1.8,
                   cat.pos = c(-30,30)
)
```



## create the list

## This represents genes called differentially expressed between pipelines for the Colitisonly vs colitis_dysplasia groups

```{r}
genelist <- list(htg = row.names(sigRes_HTG_3_Colnodysp),
          htseq = row.names(sigRes_HTSeq_3_Colnodysp)) ; head(genelist$htg) ; head(genelist$htseq)

genelist_3colnodys <- genelist

# # checking for overlap in the top 20 genes called significant and relevant
# genelist <- list(htg = row.names(sigRes_HTG_Coldysp)[1:20],
#           htseq = row.names(sigRes_HTSeq_Coldysp)[1:20]) ; head(genelist$htg) head(genelist$htseq)
```

## Plot the Venn Diagram

```{r}
ggVennDiagram(genelist)
ggVennDiagram(genelist, label_alpha = )
ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"))+
  ggplot2::scale_fill_gradient(low="darkred", high = "steelblue")

ggVennDiagram(genelist, label_alpha = 1, category.names = c("HTG pipeline genes", "HTSeq pipeline genes"),show_intersect = TRUE)

```


### to convert the Venn diagram above to a Euler diagram

# load Venn diagram package 
library("VennDiagram") 
```{r}
# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
draw.pairwise.venn(area1=length(genelist$htg), 
                   area2=length(genelist$htseq),
                   cross.area=86, 
                   category=c("HTG pipeline genes", "HTSeq pipeline genes"),
                   fill=c("red","steelblue"),
                   fontface = "bold",
                   cex = 1.8,
                   ext.text = F,
                   cat.just = rep(list(c(0.5,0.4)), 2),
                   cat.dist = rep(0.03, 2),
                   cat.fontface = rep("bold", 2),
                   cat.cex = 1.8,
                   cat.pos = c(-30,30)
)

```

## Save the results as an R image

```{r}
save.image("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/htg_hisat2-htseq_pipeline_overlap.Rdata")
```

========================================================================================
# Comparing raw gene names identified as features by the HISAT2 compared to the Mouse Ref panel by HTG   

## before filtering for the first Venn diagram, and after filtering by log2(CPM) > 0.05 for the second Venn diagram
## 30.10.2023 -->  rerun this section with the meanLog2CPM > 0.1 cut-off up from 0.05

```{r}
raw_genes <- list(HTG = row.names(counts(ddsHTG)),
                  HTSeq = row.names(counts(ddsHTSeq)))
ggVennDiagram(raw_genes)

raw_genes <- list(HTG = row.names(counts(dds_cpm_filtered_HTG1)),
                  HTSeq = row.names(counts(dds_cpm_filtered_HTSeq1)))
ggVennDiagram(raw_genes)
```


### to convert the Venn diagram above to a Euler diagram

```{r}
# load Venn diagram package 
library("VennDiagram") 

# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
draw.pairwise.venn(area1=length(raw_genes$HTG), 
                   area2=length(raw_genes$HTSeq),
                   cross.area=1610, 
                   category=c("HTG pipeline genes", "HTSeq pipeline genes"),
                   fill=c("red","steelblue"),
                   fontface = "bold",
                   cex = 1.2,
                   ext.text = F,
                   cat.just = rep(list(c(0.5,0.4)), 2),
                   cat.dist = rep(0.03, 2),
                   cat.fontface = rep("bold", 2),
                   cat.cex = 1.2,
                   cat.pos = c(-30,30)
)
```


## after filtering by log2(CPM) > 1
```{r}
# cpm_filtered_genes <- list(HTG = row.names(counts(dds_cpm_filtered_HTG)),
#                   HTSeq = row.names(counts(dds_cpm_filtered_HTSeq)))
# ggVennDiagram(cpm_filtered_genes)
```

## after filtering by raw counts > 10

```{r}
# filtered_genes <- list(HTG = row.names(counts(dds_rowSums_greaterthan_10_HTG)),
#                   HTSeq = row.names(counts(dds_rowSums_greaterthan_10_HTSeq)))
# ggVennDiagram(filtered_genes)

```


```{r}
filtered_genes <- list(HTG = row.names(counts(dds_cpm_filtered_HTG1)),
                  HTSeq = row.names(counts(dds_cpm_filtered_HTSeq1)))
ggVennDiagram(filtered_genes)
```


```{r}
paste("No. of genes uniquely present in the HTG Pipeline are", length(exclusive_HTG <- setdiff(filtered_genes$HTG,filtered_genes$HTSeq)))
(exclusive_HTG <- setdiff(filtered_genes$HTG,filtered_genes$HTSeq))

paste("No. of genes uniquely present in the HTSeq Pipeline are", length(exclusive_HTSeq <- setdiff(filtered_genes$HTSeq,filtered_genes$HTG)))
(exclusive_HTSeq <- setdiff(filtered_genes$HTSeq,filtered_genes$HTG))
```
#### note; Need to be certain that the gene names are consistent between the two pipelines. e.g., Ifna- family versus Ifna1,15,16,6, or the Ifnl2_Ifnl3 versus Ifnl2, Ifnl3... NOW WE KNOW THAT THE NAMING OF THE GENES CAN BE A CONTRIBUTING FACTOR TO THE DISPERITy BETWEEN THE TWO PIPELINES.

```{r}
# exclusive_HTG_HTSeq_filtered <- data.frame(htg=exclusive_HTSeq, 
#                                            htseq=exclusive_HTG)
# head(exclusive_HTG_HTSeq_filtered)
```


```{r}
# Make sure both vectors have the same length
(length_htg <- length(exclusive_HTG))
(length_htseq <- length(exclusive_HTSeq))

# Add NA values to the shorter vector to match lengths
exclusive_HTG_withNA <- exclusive_HTG
exclusive_HTSeq_withNA <- exclusive_HTSeq

if (length_htg < length_htseq) {
  exclusive_HTG_withNA <- c(exclusive_HTG, rep(NA, length_htseq - length_htg))
  print("first condition met")
} else if (length_htseq < length_htg) {
  exclusive_HTSeq_withNA <- c(exclusive_HTSeq, rep(NA, length_htg - length_htseq))
  print("second condition met")
}

# Now both vectors have the same length, and I can create the data frame
exclusive_HTG_HTSeq_filtered <- data.frame(htg = exclusive_HTG_withNA, htseq = exclusive_HTSeq_withNA)
head(exclusive_HTG_HTSeq_filtered)

```


```{r}
write.csv(exclusive_HTG_HTSeq_filtered,file = "C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/results/pipeline_comparisons/exclusive_HTG_HTSeq_filtered.csv")
```





## check whether these genes uniquely present in the HTG gene panel are significantly differentially regulated

## in the colitis only
```{r}
in_htg <- list("exclusive in HTG" = exclusive_HTG,"sigRes in HTG" = row.names(sigRes_HTG) )
ggVennDiagram(in_htg)

```
The above Venn Diagramm shows that 12 of the 49 genes uniquely present in HTG noCol vs Colonly samples are found to be statistically significantly DE


### to convert the Venn diagram above to a Euler diagram
```{r}
# load Venn diagram package 
library("VennDiagram") 

# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
draw.pairwise.venn(area1=length(in_htg$`exclusive in HTG`), 
                   area2=length(in_htg$`sigRes in HTG`),
                   cross.area=12, 
                   category=c("Exclusive in HTG", "Sig. and rel. in HTG"),
                   fill=c("red","steelblue"),
                   fontface = "bold",
                   cex = 1.2,
                   ext.text = F,
                   cat.just = rep(list(c(0.5,0.4)), 2),
                   cat.dist = rep(0.03, 2),
                   cat.fontface = rep("bold", 2),
                   cat.cex = 1.2,
                   cat.pos = c(-5,0)
)
```


```{r}
in_htseq <- list("exclusive in HTSeq" = exclusive_HTSeq,"sigRes in HTSeq" = row.names(sigRes_HTSeq) )
ggVennDiagram(in_htseq)
```
The above Venn diagram shows that 14 of the 56 genes uniquely present in noCol vs ColOnly samples of the HISAT2-HTSeq pipeline are called statistically significant (adj.Pvalue < 0.05) and relevant (an FDR of > 1)


### to convert the Venn diagram above to a Euler diagram
```{r}
# load Venn diagram package 
library("VennDiagram") 

# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
draw.pairwise.venn(area1=length(in_htseq$`exclusive in HTSeq`), 
                   area2=length(in_htseq$`sigRes in HTSeq`),
                   cross.area=14, 
                   category=c("Exclusive in HTSeq", "Sig. and rel. in HTSeq"),
                   fill=c("red","steelblue"),
                   fontface = "bold",
                   cex = 1.2,
                   ext.text = F,
                   cat.just = rep(list(c(0.5,0.4)), 2),
                   cat.dist = rep(0.03, 2),
                   cat.fontface = rep("bold", 2),
                   cat.cex = 1.2,
                   cat.pos = c(-5,0)
)
```





=======================

## check whether these genes uniquely present in the HTG gene panel are significantly differentially regulated

## in the col_dysp group

```{r}
in_htg <- list("exclusive in HTG" = exclusive_HTG,"sigRes in HTG" = row.names(sigRes_HTG_Coldysp) )
ggVennDiagram(in_htg)

```
The above Venn Diagramm shows that 12 of the 49 genes uniquely present in HTG noCol vs Colonly samples are found to be statistically significantly DE


### to convert the Venn diagram above to a Euler diagram
```{r}
# load Venn diagram package 
library("VennDiagram") 

# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
draw.pairwise.venn(area1=length(in_htg$`exclusive in HTG`), 
                   area2=length(in_htg$`sigRes in HTG`),
                   cross.area=13, 
                   category=c("Exclusive in HTG", "Sig. and rel. in HTG"),
                   fill=c("red","steelblue"),
                   fontface = "bold",
                   cex = 1.2,
                   ext.text = F,
                   cat.just = rep(list(c(0.5,0.4)), 2),
                   cat.dist = rep(0.03, 2),
                   cat.fontface = rep("bold", 2),
                   cat.cex = 1.2,
                   cat.pos = c(-5,0)
)
```


```{r}
in_htseq <- list("exclusive in HTSeq" = exclusive_HTSeq,"sigRes in HTSeq" = row.names(sigRes_HTSeq_Coldysp) )
ggVennDiagram(in_htseq)
```
The above Venn diagram shows that 14 of the 56 genes uniquely present in noCol vs ColOnly samples of the HISAT2-HTSeq pipeline are called statistically significant (adj.Pvalue < 0.05) and relevant (an FDR of > 1)


### to convert the Venn diagram above to a Euler diagram
```{r}
# load Venn diagram package 
library("VennDiagram") 

# move to new plotting page 
grid.newpage() 

# create pairwise Venn diagram 
draw.pairwise.venn(area1=length(in_htseq$`exclusive in HTSeq`), 
                   area2=length(in_htseq$`sigRes in HTSeq`),
                   cross.area=11, 
                   category=c("Exclusive in HTSeq", "Sig. and rel. in HTSeq"),
                   fill=c("red","steelblue"),
                   fontface = "bold",
                   cex = 1.2,
                   ext.text = F,
                   cat.just = rep(list(c(0.5,0.4)), 2),
                   cat.dist = rep(0.03, 2),
                   cat.fontface = rep("bold", 2),
                   cat.cex = 1.2,
                   cat.pos = c(-5,0)
)
```
===================================================================================

## after filtering by log2(CPM) > 0.05

###### First, create a list of the (raw) genes filtered by the above criteria and plot the Venn Diagram

```{r}
filtered_genes <- list(HTG = row.names(counts(dds_cpm_filtered_HTG1)),
                  HTSeq = row.names(counts(dds_cpm_filtered_HTSeq1)))
ggVennDiagram(filtered_genes)
```
###### Second, have a list of the unique genes in each pipeline, i.e., of the 49 and 56 genes in the HTG and HTseq pipelines, respectively


```{r}
paste("No. of genes uniquely present in the HTG Pipeline are", length(exclusive_HTG <- setdiff(filtered_genes$HTG,filtered_genes$HTSeq)))
(exclusive_HTG <- setdiff(filtered_genes$HTG,filtered_genes$HTSeq))

paste("No. of genes uniquely present in the HTSeq Pipeline are", length(exclusive_HTSeq <- setdiff(filtered_genes$HTSeq,filtered_genes$HTG)))
(exclusive_HTSeq <- setdiff(filtered_genes$HTSeq,filtered_genes$HTG))
```

#### note; Need to be certain that the gene names are consistent between the two pipelines. e.g., Ifna- family versus Ifna1,15,16,6, or the Ifnl2_Ifnl3 versus Ifnl2, Ifnl3... NOW WE KNOW THAT THE NAMING OF THE GENES CAN BE A CONTRIBUTING FACTOR TO THE DISPERITy BETWEEN THE TWO PIPELINES.

```{r}
exclusive_HTG_HTSeq_filtered <- data.frame(htg=exclusive_HTSeq, 
                                           htseq=exclusive_HTG)
head(exclusive_HTG_HTSeq_filtered)
```


```{r}
# Make sure both vectors have the same length
(length_htg <- length(exclusive_HTG))
(length_htseq <- length(exclusive_HTSeq))

# Add NA values to the shorter vector to match lengths
if (length_htg < length_htseq) {
  exclusive_HTG <- c(exclusive_HTG, rep(NA, length_htseq - length_htg))
} else if (length_htseq < length_htg) {
  exclusive_HTSeq <- c(exclusive_HTSeq, rep(NA, length_htg - length_htseq))
}

# Now both vectors have the same length, and I can create the data frame
exclusive_HTG_HTSeq_filtered <- data.frame(htg = exclusive_HTG, htseq = exclusive_HTSeq)
head(exclusive_HTG_HTSeq_filtered)

```


```{r}
write.csv(exclusive_HTG_HTSeq_filtered,file = "C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/exclusive_HTG_HTSeq_filtered.csv")
```

```{bash}
#!/bin/bash

#path=/mnt/c/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts
#path=C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts

path=C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/results/pipeline_comparisons

list1=`cut -d "," ${path}/exclusive_HTG_HTSeq_filtered.csv -f2`
list2=`cut -d "," ${path}/exclusive_HTG_HTSeq_filtered.csv -f3`

for string1 in ${list1};
do 
for string2 in ${list2};
    do
    # echo "${string1:0:4},${string2}"

    if [ "${string1:0:4}" = "${string2:0:4}" ]; then
    echo "${string1},${string2}" 
fi
    done
done > ${path}/present_in_both_P_3char.csv
# done > ${path}/trial_file.txt


for string1 in ${list1};
do 
for string2 in ${list2};
    do
    # echo "${string1:0:4},${string2}"

    if [ "${string1:0:5}" = "${string2:0:5}" ]; then
    echo "${string1},${string2}" 
fi
    done
done > ${path}/present_in_both_P_4char.csv
echo "done"
```




## check whether these genes uniquely present in the HTG gene panel are significantly differentially regulated 
### NOTE! THIS LIST NOW CONTAINS THE "NA" VALUES

```{r}
exclusive_HTG
in_htg <- list("exclusive in HTG" = exclusive_HTG,"sigRes in HTG" = row.names(sigRes_HTG) )
ggVennDiagram(in_htg)

```
The above Venn Diagramm shows that 12 of the 47 genes uniquely present in HTG noCol vs Colonly samples are found to be statistically significantly DE

### save these genes as a .csv file
```{r}
length(intersect(exclusive_HTG,row.names(sigRes_HTG)))
intersect(exclusive_HTG,row.names(sigRes_HTG))
in_htg2 <- data.frame("htg" = intersect(exclusive_HTG,row.names(sigRes_HTG))) ; in_htg2
```

```{r}
in_htseq <- list("exclusive in HTSeq" = exclusive_HTSeq,"sigRes in HTSeq" = row.names(sigRes_HTSeq) )
ggVennDiagram(in_htseq)
```
The above Venn diagram shows that 13 of the 65 genes uniquely present in noCol vs ColOnly samples of the HISAT2-HTSeq pipeline are called statistically significant (adj.Pvalue < 0.05) and relevant (an FDR of > 1)

```{r}
length(intersect(exclusive_HTSeq,row.names(sigRes_HTSeq)))
in_htseq2 <- data.frame("htseq" = intersect(exclusive_HTSeq,row.names(sigRes_HTSeq))) ; in_htseq2

```

## Write these genes diff in the colOnly in a .csv file
```{r}
# Make sure both vectors have the same length
(length_htg <- nrow(in_htg2))
(length_htseq <- nrow(in_htseq2))

# Add NA values to the shorter vector to match lengths
if (length_htg < length_htseq) {
  na_to_add <- (rep(NA, length_htseq - length_htg))
  in_htg2 <- rbind(in_htg2, na_to_add)
} else if (length_htseq < length_htg) {
  in_htseq2 <- c(in_htseq2, rep(NA, length_htg - length_htseq))
}

(length_htg <- nrow(in_htg2))
(length_htseq <- nrow(in_htseq2))

# Now both vectors have the same length, and I can create the data frame
data_frame <- data.frame("htg" <- in_htg2, "htseq" = in_htseq2)
data_frame
write.csv(data_frame, file = "C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/results/pipeline_comparisons/DEG_in_colonly_exclusively_present_in_pipelines.csv")
```

