---
title: "Jaccard funtion for the measure of similarity between two datasets"
author: "Hildah Njoroge"
date: "2023-08-16"
output: html_document
---

##### Jaccard

Jaccard Similarity = (number of observations in both sets, i.e., intersection) / (number in either set, i.e., union). This formula returns a value between 0 and 1. Values closer to 1 indicate there is a great degree of similarity between the two datasets

Step 1, define the function

```{r}
jaccard <- function(a, b) {
    intersection = length(intersect(a, b))
    union = length(a) + length(b) - intersection
    return (intersection/union)
}
```

2.  Load the data 

```{r}
load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/htg_hisat2-htseq_pipeline_overlap.Rdata")
```

3. Run the function

##### The Jaccard coeff of the noCol vs colOnly datasets
```{r}
length(genelist_colonly$htg) 
length(genelist_colonly$htseq) 

(jacc_coeff <- jaccard(genelist_colonly$htg, genelist_colonly$htseq))
paste("The Jaccard coeff of the noCol vs colOnly datasets is", round(jacc_coeff,2))
```
##### The Jaccard coeff of the noCol vs colNodys datasets
```{r}
length(genelist_colnodys$htg)
length(genelist_colnodys$htseq)

(jacc_coeff <- jaccard(genelist_colnodys$htg, genelist_colnodys$htseq))
paste("The Jaccard coeff of the noCol vs colNodys datasets is",round(jacc_coeff,2))
```


##### The Jaccard coeff of the noCol vs colDys datasets

```{r}
length(genelist_coldys$htg)
length(genelist_coldys$htseq)

(jacc_coeff <- jaccard(genelist_coldys$htg, genelist_coldys$htseq))

paste("The Jaccard coeff of the noCol vs colDys datasets is", round(jacc_coeff,2))

```

##### Jaccard distance
calculated as 1-jaccard(a,b)
This function returns distance values between 0 and 1. The greater the value, in this case, the greater the distance btn the two datasets and, consequently, the less similar they are.


================================================================================
#### The Jaccard package

##### Step 1: Install and load package
```{r}
if (!require(devtools)) {install.packages("devtools")}
# library("devtools")
if (!require("jaccard")) {devtools::install_github("ncchung/jaccard")}
```

##### Step 2: Prepare the data sets

I. Load the necessary R image to acquire the genelists
```{r}
load("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/htg_hisat2-htseq_pipeline_overlap.Rdata")
```

###### noCol vs colOnly groups
II. Prepare the data.frame with the genes present and absent indicated by 1s and 0s, respectively

```{r}
# create an empty data frame
jaccard_list_colonly <- data.frame()

for (i in 1:length(genelist_colonly$htg)) {
  # print(i)
  new_row <- data.frame(htg = 1, htseq = 0)
  jaccard_list_colonly <- rbind(jaccard_list_colonly, new_row)
}
head(jaccard_list_colonly)

## to confirm that the number of rows is correct
nrow(jaccard_list_colonly) ; length(genelist_colonly$htg)


rownames(jaccard_list_colonly) <- genelist_colonly$htg

jaccard_list_colonly2 <- data.frame()

for (i in 1:length(genelist_colonly$htseq)) {
  new_row2 <- data.frame(htg=0, htseq=1)
  jaccard_list_colonly2 <- rbind(jaccard_list_colonly2,new_row2)
}
rownames(jaccard_list_colonly2) <- genelist_colonly$htseq

jaccard_list_colonly <- rbind(jaccard_list_colonly,jaccard_list_colonly2)

## to confirm that the number of rows created is correct
nrow(jaccard_list_colonly) ; sum(length(genelist_colonly$htg), length(genelist_colonly$htseq))

head(jaccard_list_colonly)
```

```{r}
# from the Jaccard list, derive the non-unique
non_unique_rows <- rownames(jaccard_list_colonly)[duplicated(c(genelist_colonly$htg,genelist_colonly$htseq))]

```


```{r}

# create an empty data frame
jaccard_list_colonly <- data.frame()

for (i in 1:length(non_unique_rows)) {
  # print(i)
  new_row <- data.frame(htg = 1, htseq = 1)
  jaccard_list_colonly <- rbind(jaccard_list_colonly, new_row)
}
head(jaccard_list_colonly)

## to confirm that the number of rows is correct
nrow(jaccard_list_colonly) ; length(non_unique_rows)


rownames(jaccard_list_colonly) <- non_unique_rows
head(jaccard_list_colonly)


# unique in HTG pipeline
jaccard_list_colonly2 <- data.frame()
in_htg <- setdiff(genelist_colonly$htg,genelist_colonly$htseq)
for (i in 1:length(in_htg)) {
  new_row2 <- data.frame(htg=1, htseq=0)
  jaccard_list_colonly2 <- rbind(jaccard_list_colonly2,new_row2)
}
rownames(jaccard_list_colonly2) <- in_htg

jaccard_list_colonly <- rbind(jaccard_list_colonly,jaccard_list_colonly2)



# unique in HTSeq pipeline
jaccard_list_colonly3 <- data.frame()
in_htseq <- setdiff(genelist_colonly$htseq,genelist_colonly$htg)
for (i in 1:length(in_htseq)) {
  new_row3 <- data.frame(htg=0, htseq=1)
  jaccard_list_colonly3 <- rbind(jaccard_list_colonly3,new_row3)
}
rownames(jaccard_list_colonly3) <- in_htseq

jaccard_list_colonly <- rbind(jaccard_list_colonly,jaccard_list_colonly3)

## to confirm that the number of rows created is correct = 460
nrow(jaccard_list_colonly) ; sum(length(in_htg),length(in_htseq),length(non_unique_rows))

head(jaccard_list_colonly)
```

 
III. Run the Jaccard tool - first the Jaccard coeff
```{r}
#first create thethe binary vectors as input for jaccard
binary1 <- jaccard_list_colonly$htg
binary2 <- jaccard_list_colonly$htseq

jaccard::jaccard(binary1,binary2)
round(jaccard::jaccard(binary1,binary2),2)
```
IV. - the Pvalue to check the level of significance
```{r}
jaccard.test(binary1,binary2,method = "exact")
```
```{r}
bootstrap_jaccard <- jaccard.test(binary1,binary2,method = "bootstrap")
bootstrap_jaccard$pvalue

```

```{r}
jaccard.test(binary1,binary2,method = "mca")

```


###### noCol vs colnoDys groups
II. Prepare the data.frame with the genes present and absent indicated by 1s and 0s, respectively

```{r}
# create an empty data frame
jaccard_list_colnodys <- data.frame()

for (i in 1:length(genelist_colnodys$htg)) {
  # print(i)
  new_row <- data.frame(htg = 1, htseq = 0)
  jaccard_list_colnodys <- rbind(jaccard_list_colnodys, new_row)
}
head(jaccard_list_colnodys)

## to confirm that the number of rows is correct
nrow(jaccard_list_colnodys) ; length(genelist_colnodys$htg)


rownames(jaccard_list_colnodys) <- genelist_colnodys$htg

jaccard_list_colnodys2 <- data.frame()

for (i in 1:length(genelist_colnodys$htseq)) {
  new_row2 <- data.frame(htg=0, htseq=1)
  jaccard_list_colnodys2 <- rbind(jaccard_list_colnodys2,new_row2)
}
rownames(jaccard_list_colnodys2) <- genelist_colnodys$htseq

jaccard_list_colnodys <- rbind(jaccard_list_colnodys,jaccard_list_colnodys2)

## to confirm that the number of rows created is correct
nrow(jaccard_list_colnodys) ; sum(length(genelist_colnodys$htg), length(genelist_colnodys$htseq))

head(jaccard_list_colnodys)
```

```{r}
# from the Jaccard list, derive the non-unique
non_unique_rows <- rownames(jaccard_list_colnodys)[duplicated(c(genelist_colnodys$htg,genelist_colnodys$htseq))]

```


```{r}

# create an empty data frame
jaccard_list_colnodys <- data.frame()

for (i in 1:length(non_unique_rows)) {
  # print(i)
  new_row <- data.frame(htg = 1, htseq = 1)
  jaccard_list_colnodys <- rbind(jaccard_list_colnodys, new_row)
}
head(jaccard_list_colnodys)

## to confirm that the number of rows is correct
nrow(jaccard_list_colnodys) ; length(non_unique_rows)


rownames(jaccard_list_colnodys) <- non_unique_rows
head(jaccard_list_colnodys)


# unique in HTG pipeline
jaccard_list_colnodys2 <- data.frame()
in_htg <- setdiff(genelist_colnodys$htg,genelist_colnodys$htseq)
for (i in 1:length(in_htg)) {
  new_row2 <- data.frame(htg=1, htseq=0)
  jaccard_list_colnodys2 <- rbind(jaccard_list_colnodys2,new_row2)
}
rownames(jaccard_list_colnodys2) <- in_htg

jaccard_list_colnodys <- rbind(jaccard_list_colnodys,jaccard_list_colnodys2)



# unique in HTSeq pipeline
jaccard_list_colnodys3 <- data.frame()
in_htseq <- setdiff(genelist_colnodys$htseq,genelist_colnodys$htg)
for (i in 1:length(in_htseq)) {
  new_row3 <- data.frame(htg=0, htseq=1)
  jaccard_list_colnodys3 <- rbind(jaccard_list_colnodys3,new_row3)
}
rownames(jaccard_list_colnodys3) <- in_htseq

jaccard_list_colnodys <- rbind(jaccard_list_colnodys,jaccard_list_colnodys3)

## to confirm that the number of rows created is correct = 481
nrow(jaccard_list_colnodys) ; sum(length(in_htg),length(in_htseq),length(non_unique_rows))

head(jaccard_list_colnodys)
```

 
III. Run the Jaccard tool - first the Jaccard coeff
```{r}
#first create thethe binary vectors as input for jaccard
binary1 <- jaccard_list_colnodys$htg
binary2 <- jaccard_list_colnodys$htseq

jaccard::jaccard(binary1,binary2)
```

IV. - the Pvalue to check the level of significance
```{r}
jaccard.test(binary1,binary2,method = "exact")
```


```{r}
bootstrap_jaccard <- jaccard.test(binary1,binary2,method = "bootstrap")
bootstrap_jaccard$pvalue
```

```{r}
jaccard.test(binary1,binary2,method = "mca")

```



###### noCol vs coldys groups
II. Prepare the data.frame with the genes present and absent indicated by 1s and 0s, respectively

```{r}
# create an empty data frame
jaccard_list_coldys <- data.frame()

for (i in 1:length(genelist_coldys$htg)) {
  # print(i)
  new_row <- data.frame(htg = 1, htseq = 0)
  jaccard_list_coldys <- rbind(jaccard_list_coldys, new_row)
}
head(jaccard_list_coldys)

## to confirm that the number of rows is correct
nrow(jaccard_list_coldys) ; length(genelist_coldys$htg)


rownames(jaccard_list_coldys) <- genelist_coldys$htg

jaccard_list_coldys2 <- data.frame()

for (i in 1:length(genelist_coldys$htseq)) {
  new_row2 <- data.frame(htg=0, htseq=1)
  jaccard_list_coldys2 <- rbind(jaccard_list_coldys2,new_row2)
}
rownames(jaccard_list_coldys2) <- genelist_coldys$htseq

jaccard_list_coldys <- rbind(jaccard_list_coldys,jaccard_list_coldys2)

## to confirm that the number of rows created is correct
nrow(jaccard_list_coldys) ; sum(length(genelist_coldys$htg), length(genelist_coldys$htseq))

head(jaccard_list_coldys)
```

```{r}
# from the Jaccard list, derive the non-unique
non_unique_rows <- rownames(jaccard_list_coldys)[duplicated(c(genelist_coldys$htg,genelist_coldys$htseq))]

```


```{r}

# create an empty data frame
jaccard_list_coldys <- data.frame()

for (i in 1:length(non_unique_rows)) {
  # print(i)
  new_row <- data.frame(htg = 1, htseq = 1)
  jaccard_list_coldys <- rbind(jaccard_list_coldys, new_row)
}
head(jaccard_list_coldys)

## to confirm that the number of rows is correct
nrow(jaccard_list_coldys) ; length(non_unique_rows)


rownames(jaccard_list_coldys) <- non_unique_rows
head(jaccard_list_coldys)


# unique in HTG pipeline
jaccard_list_coldys2 <- data.frame()
in_htg <- setdiff(genelist_coldys$htg,genelist_coldys$htseq)
for (i in 1:length(in_htg)) {
  new_row2 <- data.frame(htg=1, htseq=0)
  jaccard_list_coldys2 <- rbind(jaccard_list_coldys2,new_row2)
}
rownames(jaccard_list_coldys2) <- in_htg

jaccard_list_coldys <- rbind(jaccard_list_coldys,jaccard_list_coldys2)



# unique in HTSeq pipeline
jaccard_list_coldys3 <- data.frame()
in_htseq <- setdiff(genelist_coldys$htseq,genelist_coldys$htg)
for (i in 1:length(in_htseq)) {
  new_row3 <- data.frame(htg=0, htseq=1)
  jaccard_list_coldys3 <- rbind(jaccard_list_coldys3,new_row3)
}
rownames(jaccard_list_coldys3) <- in_htseq

jaccard_list_coldys <- rbind(jaccard_list_coldys,jaccard_list_coldys3)

## to confirm that the number of rows created is correct
nrow(jaccard_list_coldys) ; sum(length(in_htg),length(in_htseq),length(non_unique_rows))

head(jaccard_list_coldys)
```

 
III. Run the Jaccard tool - first the Jaccard coeff
```{r}
#first create thethe binary vectors as input for jaccard
binary1 <- jaccard_list_coldys$htg
binary2 <- jaccard_list_coldys$htseq

jaccard::jaccard(binary1,binary2)
```

IV. - the Pvalue to check the level of significance
```{r}
# jaccard.test(binary1,binary2,method = "exact")
```
```{r}
bootstrap_jaccard <- jaccard.test(binary1,binary2,method = "bootstrap")
bootstrap_jaccard$pvalue
```

```{r}
# jaccard.test(binary1,binary2,method = "mca")

```

#### Save the R image
```{r}
save.image("C:/hne_files/MSc_Bioinf/MScProject-main_C/MScProject-main_10.06.23/scripts/DEG_analysis/Gene_counts/jaccard_function.RData")
```

